---
title: "Connectivity Map ligand assay data summary"
---

```{r setup, echo=F,message=F,warning=F}
library(cmapR)
library(UpSetR)
library(colorspace)
library(kableExtra)
```

# **Quantile-normalized transcriptomes from control and ligand-treated cell lines (level 3 data)**

```{r load_data_lvl3, message=F,warning=F,include=F}
if (exists("lvl3_data")) {
} else if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/lvl3_inputs.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/lvl3_inputs.RData") 
} else {
  source("lvl3_inputs.R")
}
```


## Ligands / cell lines assayed

```{r lvl3_data_summary, echo=F,message=F,warning=F,fig.height=4.5,fig.width=6.5,fig.show='hold'}
temp_ct <- sapply(ct14,function(X) 
  unique(lvl3_data@cdesc$pert_iname[lvl3_data@cdesc$cell_id == X]),
  simplify=F)
temp_ctype <- mapply(function(X,Y) sub(paste0("_",Y),"",X),
                     X=names(temp_ct),Y=ct14)
names(temp_ct) <- names(temp_ctype) <- ct14

temp_md <- data.frame(sets=ct14,
                      source.tissue=temp_ctype)
rownames(temp_md) <- ct14
temp_col <- qualitative_hcl(length(unique(temp_ctype)),palette="dark3")
names(temp_col) <- unique(temp_ctype)

upset(fromList(temp_ct),
      nsets=length(temp_ct),
      nintersects=NA,
      order.by="degree",
      set.metadata=list(data=temp_md,
                        plots=list(
                          list(type="text",
                               column="source.tissue",
                               assign=20,
                               colors=temp_col),
                          list(type="matrix_rows",
                               column="source.tissue",
                               colors=temp_col,
                               alpha=0.7))
      ),
      mb.ratio=c(0.6,0.4))

rm(list=grep("^temp",ls(),value=T))
```

In the Connectivity Map data, there are 16 ligands assayed in all 14 cell lines, and 300 ligands assayed in 8 cell lines.


```{r lvl3_data_dist, echo=F,message=F,warning=F,fig.show='hold'}
temp <- cbind(sapply(lig16,function(X)
  table(lvl3_data@cdesc$cell_id[lvl3_data@cdesc$pert_iname == X])),
  ctrl=table(lvl3_data_ctl@cdesc$cell_id))
temp <- as.data.frame(rbind(temp,Totals=colSums(temp)))
temp <- temp[c(ct14,"Total"),]
rownames(temp)[seq_along(ct14)] <- sapply(rownames(temp)[seq_along(ct14)],
                                          function(X) names(ct14)[ct14 == X])
temp$min.ctrl.ratio <- apply(temp[,colnames(temp) != "ctrl"],1,max) / temp$ctrl
kable(temp,format="html")  %>%
  kable_styling()

rm(list=grep("^temp",ls(),value=T))
```

In a subset of cell lines, ligands were tested over a variety of doses and times (and EGF was especially rigorously assayed), while in most cell lines ligands were tested at a single concentration over two time points.  Note that the number of control samples per cell type is much greater than the number of treatments.


## Overall UMAP

```{r lvl3_umap, echo=F,message=F,warning=F,fig.height=3,fig.width=9,fig.show='hold'}
if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/200630_lvl3_umap.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/200630_lvl3_umap.RData") 
} else {
  source("200326_lvl3_umap.R")
}

par(mfrow=c(1,3),mar=c(1.5,1,1,0.5),mgp=c(0,0,0))
plot(lvl3_umap$layout,pch=".",xaxt="n",yaxt="n",main="All (coloured by cell type)",xlab="UMAP1",ylab="UMAP2",
     col=qualitative_hcl(
       length(unique(lvl3_data@cdesc$cell_id)),
       palette="dark3",alpha=.7)[as.factor(c(lvl3_data@cdesc$cell_id,lvl3_data_ctl@cdesc$cell_id))])
plot(lvl3_umap$layout[colnames(lvl3_data_ctl@mat),],
     pch=".",xaxt="n",yaxt="n",main="Controls",xlab="UMAP1",ylab="UMAP2",
     xlim=range(lvl3_umap$layout[,1]),ylim=range(lvl3_umap$layout[,2]),
     col=qualitative_hcl(
       length(unique(lvl3_data@cdesc$cell_id)),
       palette="dark3",alpha=.7)[as.factor(lvl3_data_ctl@cdesc$cell_id)])
plot(lvl3_umap$layout[colnames(lvl3_data@mat),],
     pch=".",xaxt="n",yaxt="n",main="Treatments",xlab="UMAP1",ylab="UMAP2",
     xlim=range(lvl3_umap$layout[,1]),ylim=range(lvl3_umap$layout[,2]),
     col=qualitative_hcl(
       length(unique(lvl3_data@cdesc$cell_id)),
       palette="dark3",alpha=.7)[as.factor(lvl3_data@cdesc$cell_id)])

rm(list=grep("^temp",ls(),value=T))
```

Above are UMAP projections (from assayed genes) of all Connectivity Map ligand-treated and control data coloured by cell type.


## UMAP for common ligands only

Zooming in to include only the 16 ligands assayed in all cell types and their plate-matched controls (coloured in black).

```{r lvl3_umap_lig16, echo=F,message=F,warning=F,fig.height=4.5,fig.width=9,fig.show='hold'}
if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/200728_lvl3_lig16_UMAP.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/200728_lvl3_lig16_UMAP.RData") 
} else {
  source("200326_lvl3_umap.R")
}
Z <- sample(nrow(lvl3_lig16_umap$layout))
temp_lig <- factor(lvl3_data@cdesc[rownames(lvl3_lig16_umap$layout),"pert_iname"],
                   levels=lig16)
temp_lig_col <- qualitative_hcl(length(lig16),palette="dark3",alpha=0.5)[temp_lig]
temp_lig_col[is.na(temp_lig_col)] <- scales::alpha("black",0.8)
temp_lig <- as.character(temp_lig)
temp_lig[is.na(temp_lig)] <- "ctrl"
names(temp_lig) <- rownames(lvl3_lig16_umap$layout)
temp_ctrl <- sample(names(temp_lig)[temp_lig == "ctrl"])
temp_tx <- sample(names(temp_lig)[temp_lig != "ctrl"])

temp_ct <- factor(lvl3_data@cdesc[rownames(lvl3_lig16_umap$layout),"cell_id"],
                   levels=ct14)
names(temp_ct) <- rownames(lvl3_lig16_umap$layout)
temp_ct[temp_lig == "ctrl"] <- lvl3_data_ctl@cdesc[names(temp_ct)[temp_lig == "ctrl"],"cell_id"]
temp_ct_col <- qualitative_hcl(length(ct14),palette="dark3",alpha=0.5)[temp_ct]
names(temp_ct_col) <- names(temp_ct)


par(mfrow=c(1,2),mar=c(1.5,1,1,0.5),mgp=c(0,0,0))
plot(lvl3_lig16_umap$layout[Z,],pch=20,col=temp_lig_col[Z],
     xaxt="n",yaxt="n",main="Coloured by ligand",xlab="UMAP1",ylab="UMAP2")

plot(lvl3_lig16_umap$layout[temp_tx,],pch=20,col=temp_ct_col[temp_tx],
     xaxt="n",yaxt="n",xlab="UMAP1",ylab="UMAP2",
     main="Coloured by cell type (Ctrl over Tx)")
rect(par("usr")[1],par("usr")[3],par("usr")[2],par("usr")[4],col=scales::alpha("white",.7))
points(lvl3_lig16_umap$layout[temp_ctrl,],pch=20,col=temp_ct_col[temp_ctrl])


rm(list=c("Z",grep("^temp",ls(),value=T)))
```



## UMAP per cell line  

Below are UMAP projections of ligand-treated and control samples for each cell type.  Only the 16 ligands assayed in all cell types and their plate-matched controls (coloured in black) are included.

```{r lvl3_umap_ct, echo=F,message=F,warning=F,fig.height=15,fig.width=9,fig.show='hold'}
if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/200630_lvl3_ctUMAP.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/200630_lvl3_ctUMAP.RData") 
} else {
  source("200326_lvl3_umap.R")
}

par(mfrow=c(7,4),mar=c(1,1,1,0.5),mgp=c(0,0,0))
for (CT in names(lvl3_ctUMAP)) {
  if (nrow(lvl3_ctUMAP[[CT]]$layout) > 1000) {
    temp_pch <- "."; temp_cex <- 3
  } else {
    temp_pch <- 20; temp_cex <- 1
  }
  temp <- sample(nrow(lvl3_ctUMAP[[CT]]$layout))
  temp_col <- c(scales::alpha("black",0.5),
                qualitative_hcl(length(lig16),palette="dark3",alpha=0.5))
  names(temp_col) <- c("CTRL",lig16)
  temp_lig <- lvl3_data@cdesc[rownames(lvl3_ctUMAP[[CT]]$layout),"pert_iname"]
  temp_lig[is.na(temp_lig)] <- "CTRL"
  
  plot(lvl3_ctUMAP[[CT]]$layout[temp,],pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(CT,"by ligand"),
       col=temp_col[temp_lig][temp])
  
  temp_plate <- lvl3_data@cdesc[rownames(lvl3_ctUMAP[[CT]]$layout),"rna_plate"]
  temp_plate[is.na(temp_plate)] <- lvl3_data_ctl@cdesc[rownames(lvl3_ctUMAP[[CT]]$layout)[is.na(temp_plate)],"rna_plate"]
  temp_plate <- as.factor(temp_plate)
  
  plot(lvl3_ctUMAP[[CT]]$layout,pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(CT,"by plate"),
       col=qualitative_hcl(length(levels(temp_plate)),
                           palette="dark3",alpha=0.5)[temp_plate])
}

rm(list=c("CT",grep("^temp",ls(),value=T)))
```

So it seems the plate effect is strong in this data.  Might consider using the plate-based Z-scores from Cmap level 5 instead of the quantile normalized transcriptomes from level 3 that are shown here.


## UMAP per ligand treatment  

```{r lvl3_umap_lig, echo=F,message=F,warning=F,fig.height=17,fig.width=9,fig.show='hold'}
if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/200630_lvl3_ligUMAP.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/200630_lvl3_ligUMAP.RData") 
} else {
  source("200326_lvl3_umap.R")
}

par(mfrow=c(8,4),mar=c(1,1,1,0.5),mgp=c(0,0,0))
for (LIG in names(lvl3_ligUMAP)) {
  if (nrow(lvl3_ligUMAP[[LIG]]$layout) > 1000) {
    temp_pch <- "."; temp_cex <- 3
  } else {
    temp_pch <- 20; temp_cex <- 1
  }
  temp_ct <- lvl3_data@cdesc[rownames(lvl3_ligUMAP[[LIG]]$layout),"cell_id"]
  temp_ct[is.na(temp_ct)] <- lvl3_data_ctl@cdesc[rownames(lvl3_ligUMAP[[LIG]]$layout)[is.na(temp_ct)],"cell_id"]
  temp_ct <- as.factor(temp_ct)
  
  plot(lvl3_ligUMAP[[LIG]]$layout,pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(LIG,"by cell line"),
       col=qualitative_hcl(length(levels(temp_ct)),
                           palette="dark3",alpha=0.5)[temp_ct])
  
  temp_plate <- lvl3_data@cdesc[rownames(lvl3_ligUMAP[[LIG]]$layout),"rna_plate"]
  temp_plate[is.na(temp_plate)] <- lvl3_data_ctl@cdesc[rownames(lvl3_ligUMAP[[LIG]]$layout)[is.na(temp_plate)],"rna_plate"]
  temp_plate <- as.factor(temp_plate)
  
  plot(lvl3_ligUMAP[[LIG]]$layout,pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(LIG,"by plate"),
       col=qualitative_hcl(length(levels(temp_plate)),
                           palette="dark3",alpha=0.5)[temp_plate])
}

rm(list=c("LIG",grep("^temp",ls(),value=T)))
```

Looks like transcriptomes are entirely delineated by cell type from the ligand perspective.  

****

# **Robust Z-scores of differences in cell line transcriptomes following ligand treatment (level 4 data)**

```{r load_data_lvl4, message=F,warning=F,include=F}
rm(list=grep("umap",ls(),value=T))

if (exists("lvl4_data")) {
} else if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/lvl4_inputs.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/lvl4_inputs.RData") 
} else {
  source("lvl4_inputs.R")
}

if (exists("lvl4_data_all")) {
} else if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/lvl4_inputs_allgenes.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/lvl4_inputs_allgenes.RData") 
} else {
  source("lvl4_inputs_allgenes.R")
}
```

```{r lvl4_data_dist, echo=F,message=F,warning=F,fig.show='hold'}
temp <- sapply(lig16,function(X)
  table(lvl4_data@cdesc$cell_id[lvl4_data@cdesc$pert_iname == X]))
temp <- as.data.frame(rbind(temp,Total=colSums(temp)))
temp$Total <- rowSums(temp)
temp <- temp[c(ct14,"Total"),]
rownames(temp)[seq_along(ct14)] <- sapply(rownames(temp)[seq_along(ct14)],
                                          function(X) names(ct14)[ct14 == X])
kable(temp,format="html")  %>%
  kable_styling()

rm(list=grep("^temp",ls(),value=T))
```

## Overall UMAP

Last line are UMAP projections using all (including inferred) gene Z-scores.  This was included to see whether adding inferred genes improved the structure, but that doesn't seem to be the case.

```{r lvl4_umap, echo=F,message=F,warning=F,fig.height=9,fig.width=9,fig.show='hold'}
par(mfrow=c(3,3),mar=c(1.5,1,1,0.5),mgp=c(0,0,0))

if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/200630_lvl4_umap.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/200630_lvl4_umap.RData") 
} else {
  source("200326_lvl4_umap.R")
}

temp_col <- as.factor(lvl4_data@cdesc$cell_id)
temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
plot(lvl4_umap$layout,pch=".",cex=2,xaxt="n",yaxt="n",
     main="Coloured by cell line (all lig)",xlab="UMAP1",ylab="UMAP2",
     col=temp_col)

temp_col <- as.factor(lvl4_data@cdesc$pert_iname)
temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
plot(lvl4_umap$layout,pch=".",cex=2,xaxt="n",yaxt="n",
     main="Coloured by ligand (all lig)",xlab="UMAP1",ylab="UMAP2",
     col=temp_col)

temp_col <- as.factor(lvl4_data@cdesc$rna_plate)
temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
plot(lvl4_umap$layout,pch=".",cex=2,xaxt="n",yaxt="n",
     main="Coloured by plate (all lig)",xlab="UMAP1",ylab="UMAP2",
     col=temp_col)


if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/200630_lvl4_umap_lig16.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/200630_lvl4_umap_lig16.RData") 
} else {
  source("200326_lvl4_umap.R")
}

temp_col <- as.factor(lvl4_data@cdesc[rownames(lvl4_umap_lig16$layout),"cell_id"])
temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
plot(lvl4_umap_lig16$layout,pch=".",cex=3,xaxt="n",yaxt="n",
     main="Coloured by cell line (lig16)",xlab="UMAP1",ylab="UMAP2",
     col=temp_col)

temp_col <- as.factor(lvl4_data@cdesc[rownames(lvl4_umap_lig16$layout),"pert_iname"])
temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
plot(lvl4_umap_lig16$layout,pch=".",cex=3,xaxt="n",yaxt="n",
     main="Coloured by ligand (lig16)",xlab="UMAP1",ylab="UMAP2",
     col=temp_col)

temp_col <- as.factor(lvl4_data@cdesc[rownames(lvl4_umap_lig16$layout),"rna_plate"])
temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
plot(lvl4_umap_lig16$layout,pch=".",cex=3,xaxt="n",yaxt="n",
     main="Coloured by plate (lig16)",xlab="UMAP1",ylab="UMAP2",
     col=temp_col)


if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/200721_lvl4all_umap_lig16.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/200721_lvl4all_umap_lig16.RData") 
} else {
  source("200326_lvl4_umap.R")
}

temp_col <- as.factor(lvl4_data_all@cdesc[rownames(lvl4all_umap_lig16$layout),"cell_id"])
temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
plot(lvl4all_umap_lig16$layout,pch=".",cex=3,xaxt="n",yaxt="n",
     main="Coloured by cell line (lig16, all genes)",xlab="UMAP1",ylab="UMAP2",
     col=temp_col)

temp_col <- as.factor(lvl4_data_all@cdesc[rownames(lvl4all_umap_lig16$layout),"pert_iname"])
temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
plot(lvl4all_umap_lig16$layout,pch=".",cex=3,xaxt="n",yaxt="n",
     main="Coloured by ligand (lig16, all genes)",xlab="UMAP1",ylab="UMAP2",
     col=temp_col)

temp_col <- as.factor(lvl4_data_all@cdesc[rownames(lvl4all_umap_lig16$layout),"rna_plate"])
temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
plot(lvl4all_umap_lig16$layout,pch=".",cex=3,xaxt="n",yaxt="n",
     main="Coloured by plate (lig16, all genes)",xlab="UMAP1",ylab="UMAP2",
     col=temp_col)


rm(lvl4_umap,lvl4_umap_lig16,lvl4all_umap_lig16)
rm(list=grep("^temp",ls(),value=T))
```

Above are UMAP projections (from assayed genes) of all Connectivity Map ligand-treated and control data coloured by cell type.


## UMAP per cell line  

Below are UMAP projections of ligand-treated and control samples for each cell type.  Only the 16 ligands assayed in all cell types and their plate-matched controls (coloured in black) are included.

```{r lvl4_umap_ct, echo=F,message=F,warning=F,fig.height=15,fig.width=9,fig.show='hold'}
if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/200630_lvl4_ctUMAP.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/200630_lvl4_ctUMAP.RData") 
} else {
  source("200327_lvl4_umap.R")
}

par(mfrow=c(7,4),mar=c(1,1,1,0.5),mgp=c(0,0,0))
for (CT in names(lvl4_ctUMAP)) {
  if (nrow(lvl4_ctUMAP[[CT]]$layout) > 1000) {
    temp_pch <- "."; temp_cex <- 3
  } else {
    temp_pch <- 20; temp_cex <- 1
  }
  temp_col <- as.factor(lvl4_data@cdesc[rownames(lvl4_ctUMAP[[CT]]$layout),"pert_iname"])
  temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
  plot(lvl4_ctUMAP[[CT]]$layout,pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(CT,"by ligand"),
       col=temp_col)
  
  temp_col <- as.factor(lvl4_data@cdesc[rownames(lvl4_ctUMAP[[CT]]$layout),"rna_plate"])
  temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
  plot(lvl4_ctUMAP[[CT]]$layout,pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(CT,"by plate"),
       col=temp_col)
}

rm(lvl4_ctUMAP)
rm(list=c("CT",grep("^temp",ls(),value=T)))
```


## UMAP per ligand treatment  

```{r lvl4_umap_lig, echo=F,message=F,warning=F,fig.height=17,fig.width=9,fig.show='hold'}
if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/200630_lvl4_ligUMAP.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/200630_lvl4_ligUMAP.RData") 
} else {
  source("200326_lvl4_umap.R")
}

par(mfrow=c(8,4),mar=c(1,1,1,0.5),mgp=c(0,0,0))
for (LIG in names(lvl4_ligUMAP)) {
  if (nrow(lvl4_ligUMAP[[LIG]]$layout) > 1000) {
    temp_pch <- "."; temp_cex <- 3
  } else {
    temp_pch <- 20; temp_cex <- 1
  }
  temp_col <- as.factor(lvl4_data@cdesc[rownames(lvl4_ligUMAP[[LIG]]$layout),"cell_id"])
  temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
  plot(lvl4_ligUMAP[[LIG]]$layout,pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(LIG,"by cell line"),
       col=temp_col)
  
  temp_col <- as.factor(lvl4_data@cdesc[rownames(lvl4_ligUMAP[[LIG]]$layout),"rna_plate"])
  temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
  plot(lvl4_ligUMAP[[LIG]]$layout,pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(LIG,"by plate"),
       col=temp_col)
}

rm(lvl4_ligUMAP,lvl4_data,lvl4_data_ctl)
rm(list=c("LIG",grep("^temp",ls(),value=T)))
```



## UMAP per cell line  

Rerunning these using *all* genes (including inferred).

```{r lvl4_umap_ct_all, echo=F,message=F,warning=F,fig.height=15,fig.width=9,fig.show='hold'}
if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/200721_lvl4all_ctUMAP.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/200721_lvl4all_ctUMAP.RData") 
} else {
  source("200327_lvl4_umap.R")
}

par(mfrow=c(7,4),mar=c(1,1,1,0.5),mgp=c(0,0,0))
for (CT in names(lvl4_ctUMAP)) {
  if (nrow(lvl4_ctUMAP[[CT]]$layout) > 1000) {
    temp_pch <- "."; temp_cex <- 3
  } else {
    temp_pch <- 20; temp_cex <- 1
  }
  temp_col <- as.factor(lvl4_data_all@cdesc[rownames(lvl4_ctUMAP[[CT]]$layout),"pert_iname"])
  temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
  plot(lvl4_ctUMAP[[CT]]$layout,pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(CT,"by ligand"),
       col=temp_col)
  
  temp_col <- as.factor(lvl4_data_all@cdesc[rownames(lvl4_ctUMAP[[CT]]$layout),"rna_plate"])
  temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
  plot(lvl4_ctUMAP[[CT]]$layout,pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(CT,"by plate"),
       col=temp_col)
}

rm(lvl4_ctUMAP)
rm(list=c("CT",grep("^temp",ls(),value=T)))
```


## UMAP per ligand treatment  

```{r lvl4_umap_lig_all, echo=F,message=F,warning=F,fig.height=17,fig.width=9,fig.show='hold'}
if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/200721_lvl4all_ligUMAP.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/200721_lvl4all_ligUMAP.RData") 
} else {
  source("200326_lvl4_umap.R")
}

par(mfrow=c(8,4),mar=c(1,1,1,0.5),mgp=c(0,0,0))
for (LIG in names(lvl4_ligUMAP)) {
  if (nrow(lvl4_ligUMAP[[LIG]]$layout) > 1000) {
    temp_pch <- "."; temp_cex <- 3
  } else {
    temp_pch <- 20; temp_cex <- 1
  }
  temp_col <- as.factor(lvl4_data_all@cdesc[rownames(lvl4_ligUMAP[[LIG]]$layout),"cell_id"])
  temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
  plot(lvl4_ligUMAP[[LIG]]$layout,pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(LIG,"by cell line"),
       col=temp_col)
  
  temp_col <- as.factor(lvl4_data_all@cdesc[rownames(lvl4_ligUMAP[[LIG]]$layout),"rna_plate"])
  temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
  plot(lvl4_ligUMAP[[LIG]]$layout,pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(LIG,"by plate"),
       col=temp_col)
}

rm(lvl4_ligUMAP,lvl4_data_all)
rm(list=c("LIG",grep("^temp",ls(),value=T)))
```


****  


# **Replicate-collapsed robust Z-scores of differences in cell line transcriptomes following ligand treatment (level 5 data)**

```{r load_data_lvl5, message=F, warning=F, include=F}
if (exists("lvl5_data")) {
} else if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/lvl5_inputs.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/lvl5_inputs.RData") 
} else {
  source("lvl5_inputs.R")
}
```


# Ligands / cell lines assayed

```{r lvl5_data_summary, echo=F, message=F, warning=F, fig.height=5, fig.width=8, fig.show='hold'}
temp_ct <- sapply(ct14,function(X) 
  unique(lvl5_data@cdesc$pert_iname[lvl5_data@cdesc$cell_id == X]),
  simplify=F)
temp_ctype <- mapply(function(X,Y) sub(paste0("_",Y),"",X),
                     X=names(temp_ct),Y=ct14)
names(temp_ct) <- names(temp_ctype) <- ct14

temp_md <- data.frame(sets=ct14,
                      source.tissue=temp_ctype)
rownames(temp_md) <- ct14
temp_col <- qualitative_hcl(length(unique(temp_ctype)),palette="dark3")
names(temp_col) <- unique(temp_ctype)

upset(fromList(temp_ct),
      nsets=length(temp_ct),
      nintersects=NA,
      order.by="degree",
      set.metadata=list(data=temp_md,
                        plots=list(
                          list(type="text",
                               column="source.tissue",
                               assign=15,
                               colors=temp_col),
                          list(type="matrix_rows",
                               column="source.tissue",
                               colors=temp_col,
                               alpha=0.7))
      ))

rm(list=grep("^temp",ls(),value=T))
```

In the Connectivity Map data, there are 16 ligands assayed in all 14 cell lines, and 300 ligands assayed in 8 cell lines.

```{r lvl5_data_dist, echo=F, message=F, warning=F, fig.show='hold'}
temp <- sapply(lig16,function(X)
  table(lvl5_data@cdesc$cell_id[lvl5_data@cdesc$pert_iname == X]))
temp <- as.data.frame(rbind(temp,Totals=colSums(temp)))
temp <- temp[c(ct14,"Total"),]
rownames(temp)[seq_along(ct14)] <- sapply(rownames(temp)[seq_along(ct14)],
                                          function(X) names(ct14)[ct14 == X])

kable(temp,format="html")  %>%
  kable_styling()

rm(list=grep("^temp",ls(),value=T))
```

In a subset of cell lines, ligands were tested over a variety of doses and times (and EGF was especially rigorously assayed), while in most cell lines ligands were tested at a single concentration over two time points.  Note that these data (level 5) are replicate-aggregated Z-scores, so each reported assay represents multiple replicates.

## Overall UMAP

```{r lvl5_umap, echo=F,message=F,warning=F,fig.height=6,fig.width=9,fig.show='hold'}
par(mfrow=c(2,3),mar=c(1.5,1,1,0.5),mgp=c(0,0,0))

if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/200701_lvl5_umap.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/200701_lvl5_umap.RData") 
} else {
  source("200701_lvl5_umap.R")
}

temp_col <- as.factor(lvl5_data@cdesc$cell_id)
temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
plot(lvl5_umap$layout,pch=".",xaxt="n",yaxt="n",
     main="Coloured by cell line",xlab="UMAP1",ylab="UMAP2",
     col=temp_col)

temp_col <- as.factor(lvl5_data@cdesc$pert_iname)
temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
plot(lvl5_umap$layout,pch=".",xaxt="n",yaxt="n",
     main="Coloured by ligand (all)",xlab="UMAP1",ylab="UMAP2",
     col=temp_col)

temp_col <- as.factor(lvl5_data@cdesc$pert_time)
temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
plot(lvl5_umap$layout,pch=".",xaxt="n",yaxt="n",
     main="Coloured by duration (correlates with plate)",xlab="UMAP1",ylab="UMAP2",
     col=temp_col)


if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/200701_lvl5_umap_lig16.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/200701_lvl5_umap_lig16.RData") 
} else {
  source("200701_lvl5_umap.R")
}

temp_col <- as.factor(lvl5_data@cdesc[rownames(lvl5_umap_lig16$layout),"cell_id"])
temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
plot(lvl5_umap_lig16$layout,pch=20,xaxt="n",yaxt="n",
     main="Coloured by cell line",xlab="UMAP1",ylab="UMAP2",
     col=temp_col)

temp_col <- as.factor(lvl5_data@cdesc[rownames(lvl5_umap_lig16$layout),"pert_iname"])
temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
plot(lvl5_umap_lig16$layout,pch=20,xaxt="n",yaxt="n",
     main="Coloured by ligand (common only)",xlab="UMAP1",ylab="UMAP2",
     col=temp_col)

temp_col <- as.factor(lvl5_data@cdesc[rownames(lvl5_umap_lig16$layout),"pert_time"])
temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
plot(lvl5_umap_lig16$layout,pch=20,xaxt="n",yaxt="n",
     main="Coloured by duration (correlates with plate)",xlab="UMAP1",ylab="UMAP2",
     col=temp_col)


rm(list=grep("^temp",ls(),value=T))
```

Above are UMAP projections (from assayed genes) of all Connectivity Map ligand-treated and control data coloured by cell type.


## UMAP per cell line  

Below are UMAP projections of ligand-treated and control samples for each cell type.  Only the 16 ligands assayed in all cell types and their plate-matched controls (coloured in black) are included.

```{r lvl5_umap_ct, echo=F,message=F,warning=F,fig.height=15,fig.width=9,fig.show='hold'}
if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/200701_lvl5_ctUMAP.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/200701_lvl5_ctUMAP.RData") 
} else {
  source("200327_lvl5_umap.R")
}

par(mfrow=c(7,4),mar=c(1,1,1,0.5),mgp=c(0,0,0))
for (CT in names(lvl5_ctUMAP)) {
  if (nrow(lvl5_ctUMAP[[CT]]$layout) > 1000) {
    temp_pch <- "."; temp_cex <- 3
  } else {
    temp_pch <- 20; temp_cex <- 1
  }
  temp_col <- as.factor(lvl5_data@cdesc[rownames(lvl5_ctUMAP[[CT]]$layout),"pert_iname"])
  temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
  plot(lvl5_ctUMAP[[CT]]$layout,pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(CT,"by ligand"),
       col=temp_col)
  
  temp_col <- as.factor(lvl5_data@cdesc[rownames(lvl5_ctUMAP[[CT]]$layout),"pert_time"])
  temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
  plot(lvl5_ctUMAP[[CT]]$layout,pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(CT,"by duration (plate)"),
       col=temp_col)

}

rm(list=c("CT",grep("^temp",ls(),value=T)))
```


## UMAP per ligand treatment  

```{r lvl5_umap_lig, echo=F,message=F,warning=F,fig.height=17,fig.width=9,fig.show='hold'}
if (file.exists("~/Dropbox/GDB_archive/CMapCorr_files/200701_lvl5_ligUMAP.RData")) {
  load("~/Dropbox/GDB_archive/CMapCorr_files/200701_lvl5_ligUMAP.RData") 
} else {
  source("200701_lvl5_umap.R")
}

par(mfrow=c(8,4),mar=c(1,1,1,0.5),mgp=c(0,0,0))
for (LIG in names(lvl5_ligUMAP)) {
  if (nrow(lvl5_ligUMAP[[LIG]]$layout) > 1000) {
    temp_pch <- "."; temp_cex <- 3
  } else {
    temp_pch <- 20; temp_cex <- 1
  }
  temp_col <- as.factor(lvl5_data@cdesc[rownames(lvl5_ligUMAP[[LIG]]$layout),"cell_id"])
  temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
  plot(lvl5_ligUMAP[[LIG]]$layout,pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(LIG,"by cell line"),
       col=temp_col)

  temp_col <- as.factor(lvl5_data@cdesc[rownames(lvl5_ligUMAP[[LIG]]$layout),"pert_time"])
  temp_col <- qualitative_hcl(length(levels(temp_col)),palette="dark3",alpha=.7)[temp_col]
  plot(lvl5_ligUMAP[[LIG]]$layout,pch=temp_pch,cex=temp_cex,xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste(LIG,"by duration (plate)"),
       col=temp_col)
}

rm(list=c("LIG",grep("^temp",ls(),value=T)))
```

