---
title: "Cmap data summary"
---

```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
library(cmapR)
library(UpSetR)
library(ranger)
library(colorspace)
library(kableExtra)
```

```{r load_data_lvl3, message=FALSE, warning=FALSE, include=FALSE}
if (file.exists("../CMapCorr_files/lvl3_inputs.RData") & !exists("lvl3_data")) {
  load("../CMapCorr_files/lvl3_inputs.RData") 
} else {
  source("lvl3_inputs.R")
}
```


# Ligand transcriptome perturbations from Connectivity Map

## Ligands / cell lines assayed

```{r data_summary, echo=FALSE, message=FALSE, warning=FALSE, fig.height=5, fig.width=8, fig.show='hold'}
temp_ct <- sapply(unique(lvl3_data@cdesc$cell_id),function(X) 
  unique(lvl3_data@cdesc$pert_iname[lvl3_data@cdesc$cell_id == X]),
  simplify=F)
temp_ctype <- apply(cell_info[unique(lvl3_data@cdesc$cell_id),c("primary_site","sample_type")],1,
                    function(X) paste(X,collapse=" "))
temp_md <- data.frame(sets=names(temp_ctype),source.tissue=temp_ctype)
temp_col <- qualitative_hcl(length(unique(temp_ctype)),palette="dark3")
names(temp_col) <- unique(temp_md$source.tissue)

upset(fromList(temp_ct),
      nsets=length(temp_ct),
      nintersects=NA,
      order.by="degree",
      set.metadata=list(data=temp_md,
                        plots=list(
                          list(type="text",
                               column="source.tissue",
                               assign=15,
                               colors=temp_col),
                          list(type="matrix_rows",
                               column="source.tissue",
                               colors=temp_col,
                               alpha=0.7))
      ))

rm(list=grep("^temp",ls(),value=T))
```
In the Connectivity Map data, there are 15 ligands assayed in all 14 cell lines, and 300 ligands assayed in 8 cell lines.


## Distribution of samples between cell types and ligands

```{r lvl3_pca, echo=FALSE, message=FALSE, warning=FALSE, fig.height=9,fig.width=9, fig.show='hold'}
if (file.exists("../CMapCorr_files/200326_lvl3pca.RData") & !exists("lvl3pca")) {
  load("../CMapCorr_files/200326_lvl3pca.RData") 
} else {
  source("200326_lvl3_pca_umap.R")
}

par(mfrow=c(3,3),mar=c(1.5,1,1,0.5),mgp=c(0,0,0))
for (X in list(c(1,2),c(1,3),c(3,2))) {
  plot(lvl3pca$x[,X],pch=".",xaxt="n",yaxt="n",xaxt="n",yaxt="n",main="All",
       col=qualitative_hcl(
         length(unique(lvl3_data@cdesc$cell_id)),
         palette="dark3",alpha=.7)[as.factor(c(lvl3_data@cdesc$cell_id,lvl3_data_ctl@cdesc$cell_id))])
}
for (X in list(c(1,2),c(1,3),c(3,2))) {
  plot(lvl3pca$x[colnames(lvl3_data_ctl@mat),X],pch=".",xaxt="n",yaxt="n",main="Controls",
       col=qualitative_hcl(
         length(unique(lvl3_data@cdesc$cell_id)),
         palette="dark3",alpha=.7)[as.factor(lvl3_data_ctl@cdesc$cell_id)])
}
for (X in list(c(1,2),c(1,3),c(3,2))) {
  plot(lvl3pca$x[colnames(lvl3_data@mat),X],pch=".",xaxt="n",yaxt="n",main="Treatments",
       col=qualitative_hcl(
         length(unique(lvl3_data@cdesc$cell_id)),
         palette="dark3",alpha=.7)[as.factor(lvl3_data@cdesc$cell_id)])
}

rm(list=grep("^temp",ls(),value=T))
```

```{r lvl3_umap, echo=FALSE, message=FALSE, warning=FALSE, fig.height=6, fig.width=9, fig.show='hold'}
if (file.exists("../CMapCorr_files/200326_lvl3umap.RData") & !exists("lvl3umap")) {
  load("../CMapCorr_files/200326_lvl3umap.RData") 
} else {
  source("200326_lvl3_pca_umap.R")
}

par(mfrow=c(2,3),mar=c(1.5,1,1,0.5),mgp=c(0,0,0))
plot(lvl3umap$layout,pch=".",xaxt="n",yaxt="n",main="All",xlab="UMAP1",ylab="UMAP2",
     col=qualitative_hcl(
       length(unique(lvl3_data@cdesc$cell_id)),
       palette="dark3",alpha=.7)[as.factor(c(lvl3_data@cdesc$cell_id,lvl3_data_ctl@cdesc$cell_id))])
plot(lvl3umap$layout[colnames(lvl3_data_ctl@mat),],
     pch=".",xaxt="n",yaxt="n",main="Controls",xlab="UMAP1",ylab="UMAP2",
     xlim=range(lvl3umap$layout[,1]),ylim=range(lvl3umap$layout[,2]),
     col=qualitative_hcl(
       length(unique(lvl3_data@cdesc$cell_id)),
       palette="dark3",alpha=.7)[as.factor(lvl3_data_ctl@cdesc$cell_id)])
plot(lvl3umap$layout[colnames(lvl3_data@mat),],
     pch=".",xaxt="n",yaxt="n",main="Treatments",xlab="UMAP1",ylab="UMAP2",
     xlim=range(lvl3umap$layout[,1]),ylim=range(lvl3umap$layout[,2]),
     col=qualitative_hcl(
       length(unique(lvl3_data@cdesc$cell_id)),
       palette="dark3",alpha=.7)[as.factor(lvl3_data@cdesc$cell_id)])

if (file.exists("../CMapCorr_files/200326_lvl3umap_raw.RData") & !exists("lvl3umap_raw")) {
  load("../CMapCorr_files/200326_lvl3umap_raw.RData") 
} else {
  source("200326_lvl3_pca_umap.R")
}

plot(lvl3umap_raw$layout,pch=".",xaxt="n",yaxt="n",main="All",xlab="UMAP1",ylab="UMAP2",
     col=qualitative_hcl(
       length(unique(lvl3_data@cdesc$cell_id)),
       palette="dark3",alpha=.7)[as.factor(c(lvl3_data@cdesc$cell_id,lvl3_data_ctl@cdesc$cell_id))])
plot(lvl3umap_raw$layout[colnames(lvl3_data_ctl@mat),],
     pch=".",xaxt="n",yaxt="n",main="Controls",xlab="UMAP1",ylab="UMAP2",
     xlim=range(lvl3umap_raw$layout[,1]),ylim=range(lvl3umap_raw$layout[,2]),
     col=qualitative_hcl(
       length(unique(lvl3_data@cdesc$cell_id)),
       palette="dark3",alpha=.7)[as.factor(lvl3_data_ctl@cdesc$cell_id)])
plot(lvl3umap_raw$layout[colnames(lvl3_data@mat),],
     pch=".",xaxt="n",yaxt="n",main="Treatments",xlab="UMAP1",ylab="UMAP2",
     xlim=range(lvl3umap_raw$layout[,1]),ylim=range(lvl3umap_raw$layout[,2]),
     col=qualitative_hcl(
       length(unique(lvl3_data@cdesc$cell_id)),
       palette="dark3",alpha=.7)[as.factor(lvl3_data@cdesc$cell_id)])

rm(list=grep("^temp",ls(),value=T))
```

Above are PCA and UMAP projections of all Connectivity Map ligand-treated and control data coloured by cell type.

```{r data_dist, echo=FALSE, message=FALSE, warning=FALSE, fig.show='hold'}
temp <- as.data.frame(cbind(sapply(lig15,function(X)
  table(lvl3_data@cdesc$cell_id[lvl3_data@cdesc$pert_iname == X])),
  ctrl=table(lvl3_data_ctl@cdesc$cell_id)))
rownames(temp) <- sapply(rownames(temp),function(X) names(ct14)[ct14 == X])
temp$min.ctrl.ratio <- apply(temp[,colnames(temp) != "ctrl"],1,max) / temp$ctrl
kable(temp,format="html")  %>%
  kable_styling() %>%
  scroll_box(width = "9in", height = "4in")
# ct_strat <- ct14[rownames(temp)[temp$min.ctrl.ratio < 0.4]]

rm(list=grep("^temp",ls(),value=T))
```

Below are UMAP projections of selected cell types that represent patterns of data distribution between controls and treatments that we see in the CMap data. BT20 is a breast tumour line where very few ligands were tested, but they were tested in depth, and thus there are more treated samples (as a whole) than controls.  VCAP is a prostate tumour line where many ligands were minimally tested, with many controls.  The MCF7 breast tumour line is unique in that many ligands were tested in depth, so as a whole there is a balance between treated and control samples.  

```{r lvl3_umap_ct, echo=FALSE, message=FALSE, warning=FALSE, fig.height=18, fig.width=9, fig.show='hold'}
if (file.exists("../CMapCorr_files/200326_ctUMAP.RData") & !exists("ctUMAP")) {
  load("../CMapCorr_files/200326_ctUMAP.RData") 
} else {
  source("200326_lvl3_pca_umap.R")
}
par(mfrow=c(6,3),mar=c(1,1,1,0.5),mgp=c(0,0,0))
for (CT in names(ctUMAP)) {
  plot(ctUMAP[[CT]]$layout,pch=".",xaxt="n",yaxt="n",
       xlab="UMAP1",ylab="UMAP2",main=paste0(CT,": All"),
       col=scales::alpha(c("red","blue"),0.7)[rownames(ctUMAP[[CT]]$layout) %in% 
                                                colnames(lvl3_data@mat) + 1])
  legend("topright",bty="n",pch=20,col=c("red","blue"),cex=0.8,legend=c("Ctl","Tx"))
  
  temp_facts <- list(
    "Controls by solute & time"=as.factor(apply(lvl3_data_ctl@cdesc[lvl3_data_ctl@cdesc$cell_id == CT,
                                                                    c("pert_iname","pert_time")],1,
                                                function(Z) paste(Z,collapse="_"))),
    "Controls by plate"=as.factor(lvl3_data_ctl@cdesc[lvl3_data_ctl@cdesc$cell_id == CT,"rna_plate"]),
    "Treatments by ligand"=as.factor(lvl3_data@cdesc[lvl3_data@cdesc$cell_id == CT,"pert_iname"]),
    "Treatments by dose & time"=as.factor(apply(lvl3_data@cdesc[lvl3_data@cdesc$cell_id == CT,
                                                                c("pert_dose","pert_time")],1,
                                                function(Z) paste(Z,collapse="_"))),
    "Treatments by plate"=as.factor(lvl3_data@cdesc[lvl3_data@cdesc$cell_id == CT,"rna_plate"])
  )
  names(temp_facts[[2]]) <- rownames(lvl3_data_ctl@cdesc[lvl3_data_ctl@cdesc$cell_id == CT,])
  names(temp_facts[[3]]) <- names(temp_facts[[5]]) <- rownames(lvl3_data@cdesc[lvl3_data@cdesc$cell_id == CT,])

  for (X in names(temp_facts)) {
    plot(ctUMAP[[CT]]$layout[names(temp_facts[[X]]),],
         pch=20,xaxt="n",yaxt="n",xlab="UMAP1",ylab="UMAP2",
         main=paste0(CT,": ",X),
         xlim=range(ctUMAP[[CT]]$layout[,1]),ylim=range(ctUMAP[[CT]]$layout[,2]),
         col=qualitative_hcl(length(levels(temp_facts[[X]])),
                                         palette="dark3",alpha=.7)[temp_facts[[X]]])
  }
}
rm(X,CT)
rm(list=grep("^temp",ls(),value=T))
```

So it seems the plate effect is strong in this data.  Might consider using the plate-based Z-scores from Cmap level 5 instead of the quantile normalized transcriptomes from level 3 that are shown here.  For now, continuing with level 3 data.

****
