---
title: "GSEA"
---
  
```{r setup, echo=F,message=F,warning=F}
library(scales)
library(colorspace)
source("~/Dropbox/GDB/line2user.R")
par(mar=c(3,3,2,1),mgp=2:0)
.PAR <- par(no.readonly=T)
knitr::opts_chunk$set(
  echo=F,
  message=F,
  warning=F
)
```

# Mean signed FDR (downsampled)  

Since number of samples being averaged was causing artifacts, downsampled to averaging 5 randomly selected samples from each sample set.  Repeated enough times to cover the largest set once.

## Overall and within-set comparisons

Left: FDR-correction for all sample sets (ie. cell types / ligands) together, per downsample.  
Right: FDR-correction per sample set, per downsample.  

```{r overall, fig.height=9,fig.width=9,warning=FALSE}
allCuts <- seq(4,0,by=-0.01)

par(mfrow=c(2,2),mar=c(3,3,2,1),mgp=2:0)
for (GSEA in c("REACT","GOBP")) {
  load(paste0("~/Dropbox/GDB_archive/CMapCorr_files/pfdrGSEA_",GSEA,"_DS.RData"))
  
  temp_max <- allCuts[min(sapply(list(NQall_CT,NQall_LIG,NQall_LIGCT),
                                 function(X) which(colMeans(X) > 0)[1] - 1))]
  plot(NA,NA,xlim=c(0,temp_max),xaxt="n",ylim=c(0,1),main=GSEA,cex.lab=0.9,
       xlab="Threshold of FDR-corrected P(mean signed GSEA FDR) by chance",
       ylab="Mean proportion of FDR \u2264 x for sample set averages")
  axis(side=1,labels=c(.0001,.001,.01,.05,.1,.5,1),
       at=-log10(c(.0001,.001,.01,.05,.1,.5,1)))
  temp <- apply(NQall_CT[sample(nrow(NQall_CT),nrow(NQall_LIGCT)),],1,function(X)
    lines(allCuts,X,lwd=1,col=qualitative_hcl(3,palette="set2",alpha=0.5)[1]))
  temp <- apply(NQall_LIG[sample(nrow(NQall_LIG),nrow(NQall_LIGCT)),],1,function(X)
    lines(allCuts,X,lwd=1,col=qualitative_hcl(3,palette="set2",alpha=0.5)[2]))
  temp <- apply(NQall_LIGCT,1,function(X)
    lines(allCuts,X,lwd=1,col=qualitative_hcl(3,palette="set2",alpha=0.5)[3]))
  lines(allCuts,colMeans(NQall_CT),lwd=2,lty=3,col=qualitative_hcl(3,palette="dark3")[1])
  lines(allCuts,colMeans(NQall_LIG),lwd=2,lty=3,col=qualitative_hcl(3,palette="dark3")[2])
  lines(allCuts,colMeans(NQall_LIGCT),lwd=2,lty=3,col=qualitative_hcl(3,palette="dark3")[3])
  
  temp_c2l <- ks.test(colMeans(NQall_CT),colMeans(NQall_LIG))$p.value
  temp_l2lc <- ks.test(colMeans(NQall_LIG),colMeans(NQall_LIGCT))$p.value
  legend("topright",bty="n",lwd=c(1,NA,1,NA,1),
         col=as.vector(sapply(qualitative_hcl(3,palette="dark3"),function(X) c(X,NA)))[1:5],
         legend=c("Cell type",
                  paste("KS p =",signif(temp_c2l,2)),
                  "Ligand",
                  paste("KS p =",signif(temp_l2lc,2)),
                  "Lig / CT"))
  

  plot(NA,NA,xlim=c(0,temp_max),xaxt="n",ylim=c(0,1),main=GSEA,cex.lab=0.9,
       xlab="Threshold of FDR-corrected P(mean signed GSEA FDR) by chance",
       ylab="Mean proportion of FDR \u2264 x for sample set averages")
  axis(side=1,labels=c(.0001,.001,.01,.05,.1,.5,1),
       at=-log10(c(.0001,.001,.01,.05,.1,.5,1)))
  temp <- apply(NQper_CT[sample(nrow(NQper_CT),nrow(NQper_LIGCT)),],1,function(X)
    lines(allCuts,X,lwd=1,col=qualitative_hcl(3,palette="set2",alpha=0.5)[1]))
  temp <- apply(NQper_LIG[sample(nrow(NQper_LIG),nrow(NQper_LIGCT)),],1,function(X)
    lines(allCuts,X,lwd=1,col=qualitative_hcl(3,palette="set2",alpha=0.5)[2]))
  temp <- apply(NQper_LIGCT,1,function(X)
    lines(allCuts,X,lwd=1,col=qualitative_hcl(3,palette="set2",alpha=0.5)[3]))
  lines(allCuts,colMeans(NQper_CT),lwd=2,lty=3,col=qualitative_hcl(3,palette="dark3")[1])
  lines(allCuts,colMeans(NQper_LIG),lwd=2,lty=3,col=qualitative_hcl(3,palette="dark3")[2])
  lines(allCuts,colMeans(NQper_LIGCT),lwd=2,lty=3,col=qualitative_hcl(3,palette="dark3")[3])
  
  temp_c2l <- ks.test(colMeans(NQper_CT),colMeans(NQper_LIG))$p.value
  temp_l2lc <- ks.test(colMeans(NQper_LIG),colMeans(NQper_LIGCT))$p.value
  legend("topright",bty="n",lwd=c(1,NA,1,NA,1),
         col=as.vector(sapply(qualitative_hcl(3,palette="dark3"),function(X) c(X,NA)))[1:5],
         legend=c("Cell type",
                  paste("KS p =",signif(temp_c2l,2)),
                  "Ligand",
                  paste("KS p =",signif(temp_l2lc,2)),
                  "Lig / CT"))
}
rm(list=grep("^temp",ls(),value=T))
```



## Cell line  

```{r CT, fig.height=3,fig.width=9}
FDRcut <- 0.9

tempMAT <- apply(
  sapply(seq_along(Qper_CT),function(X) 
    -log10(Qper_CT[[X]]) * sign(Mscore_CT[[X]]),
    simplify="array"),
  MARGIN=c(1,2),mean)
tempMAT <- tempMAT[apply(tempMAT,1,function(X) any(abs(X) >= -log10(FDRcut))),]
temp_lim <- max(abs(tempMAT))
temp_hSample <- hclust(dist(t(tempMAT)),method="ward.D2") 
temp_hSet <- hclust(dist(tempMAT),method="ward.D2")
tempMAT <- tempMAT[temp_hSet$order,temp_hSample$order]

par(mar=c(2,6,0.5,0.5))
image(z=tempMAT,x=1:nrow(tempMAT),y=1:ncol(tempMAT),
      col=diverging_hcl(1000,palette="Blue-Red3"),
      breaks=seq(-temp_lim,temp_lim,length.out=1001),
      xaxt="n",yaxt="n",xlab=NA,ylab=NA)
mtext(colnames(tempMAT),side=2,line=0.1,at=1:ncol(tempMAT),las=2,cex=0.7)
mtext(paste(GSEA,"gene set enrichment"),side=1)
mtext(paste0("|signed FDR| \u2264 ",FDRcut * 100,"%"),side=1,line=1)

rm(list=grep("^temp",ls(),value=T))
```

## Ligand  

```{r LIG, fig.height=3,fig.width=9}
FDRcut <- 0.1

tempMAT <- apply(
  sapply(seq_along(Qper_LIG),function(X) 
    -log10(Qper_LIG[[X]]) * sign(Mscore_LIG[[X]]),
    simplify="array"),
  MARGIN=c(1,2),mean)
tempMAT <- tempMAT[,apply(tempMAT,2,function(X) any(abs(X) >= -log10(FDRcut)))]
tempMAT <- tempMAT[apply(tempMAT,1,function(X) any(abs(X) >= -log10(FDRcut))),]
temp_lim <- max(abs(tempMAT))
temp_hSample <- hclust(dist(t(tempMAT)),method="ward.D2") 
temp_hSet <- hclust(dist(tempMAT),method="ward.D2")
tempMAT <- tempMAT[temp_hSet$order,temp_hSample$order]

par(mar=c(2,4,0.5,0.5))
image(z=tempMAT,x=1:nrow(tempMAT),y=1:ncol(tempMAT),
      col=diverging_hcl(1000,palette="Blue-Red3"),
      breaks=seq(-temp_lim,temp_lim,length.out=1001),
      xaxt="n",yaxt="n",xlab=NA,ylab=NA)
mtext(colnames(tempMAT),side=2,line=0.1,at=1:ncol(tempMAT),las=2,cex=0.7)
mtext(paste0("Ligands (",ncol(tempMAT),"/",length(Mscore_LIG),")"),side=2,line=3)
mtext(paste(GSEA,"gene set enrichment"),side=1)
mtext(paste0("|signed FDR| \u2264 ",FDRcut * 100,"%"),side=1,line=1)

rm(list=grep("^temp",ls(),value=T))
```

## Ligand per cell line  

```{r LIGCT, fig.height=6,fig.width=9}
tempMAT <- apply(
  sapply(seq_along(Qper_LIGCT),function(X) 
    -log10(Qper_LIGCT[[X]]) * sign(Mscore_LIGCT[[X]]),
    simplify="array"),
  MARGIN=c(1,2),mean)
tempMAT <- tempMAT[,apply(tempMAT,2,function(X) any(abs(X) >= -log10(FDRcut)))]
tempMAT <- tempMAT[apply(tempMAT,1,function(X) any(abs(X) >= -log10(FDRcut))),]
temp_lim <- max(abs(tempMAT))
temp_hSample <- hclust(dist(t(tempMAT)),method="ward.D2") 
temp_hSet <- hclust(dist(tempMAT),method="ward.D2")
tempMAT <- tempMAT[temp_hSet$order,temp_hSample$order]

par(mar=c(2,1.5,0.5,0.5))
image(z=tempMAT,x=1:nrow(tempMAT),y=1:ncol(tempMAT),
      col=diverging_hcl(1000,palette="Blue-Red3"),
      breaks=seq(-temp_lim,temp_lim,length.out=1001),
      xaxt="n",yaxt="n",xlab=NA,ylab=NA)
# mtext(colnames(tempMAT),side=2,line=0.1,at=1:ncol(tempMAT),las=2,cex=0.7)
mtext(paste0("Ligands per cell line (",
             length(unique(sub("^.+_","",colnames(tempMAT)))),
             "/295 unique ligands)"),
      side=2,line=0.1)
mtext(paste(GSEA,"gene set enrichment"),side=1)
mtext(paste0("|signed FDR| \u2264 ",FDRcut * 100,"%"),side=1,line=1)

rm(list=grep("^temp",ls(),value=T))
```


