---
title: "GSEA"
---
  
```{r setup, echo=F,message=F,warning=F}
library(scales)
library(colorspace)
source("~/Dropbox/GDB/line2user.R")
par(mar=c(3,3,2,1),mgp=2:0)
.PAR <- par(no.readonly=T)
knitr::opts_chunk$set(
  echo=F,
  message=F,
  warning=F
)
```

# Enrichment score correlation

```{r FX}
CORR_BXP <- function(Z) {
  files <- grep(paste0("_",Z,".RData$"),
                list.files("~/Dropbox/GDB_archive/CMapCorr_files/",pattern="^corrGSEA_[A-Z]"),
                value=T)
  for (GSEA in unique(sapply(strsplit(files,"_"),"[[",2))) {
    load(paste0("~/Dropbox/GDB_archive/CMapCorr_files/corrGSEA_",GSEA,"_",Z,".RData"))
    
    temp_ds <- GSEA
    
    if (Z %in% c("ligct","rep")) {
      temp_lig <- as.factor(sub("_[A-Za-z0-9_]+$","",names(CORS)))
      temp_ord1 <- tapply(CORS,temp_lig,function(X) names(X)[order(sapply(X,median))])
      temp_ord2 <- order(tapply(CORS,temp_lig,function(X) median(unlist(X))))
      temp_order <- temp_ord1[temp_ord2]
      CORS <- CORS[unlist(temp_order)]
      temp_border <- qualitative_hcl(length(temp_order),palette="dark3")
      temp_border <- unlist(mapply(rep,
                                   x=temp_border,
                                   times=sapply(temp_order,length),
                                   SIMPLIFY=F))
      temp_col <- alpha(temp_border,0.5)
      temp_bot <- switch(1 + (length(temp_order) < 20),1.5,4)
    } else {
      temp_order <- order(sapply(CORS,median))
      CORS <- CORS[temp_order]
      temp_border <- qualitative_hcl(length(CORS),palette="dark3")
      temp_col <- qualitative_hcl(length(CORS),palette="dark3",alpha=0.5)
      temp_bot <- switch(1 + (length(CORS) < 20),1.5,4)
    }
    if (length(CORS) > 300) {
      temp_border <- alpha(temp_border,0.5)
      temp_col <- alpha(temp_col,0.2)
    }      
    par(mar=c(temp_bot,4,2,1),mgp=2:0)
    boxplot(CORS,outline=T,pch=".",xaxt="n",yaxs="i",ylim=c(-1,1),
            border=temp_border,col=temp_col,
            ylab=paste("Pairwise Spearman correlation of",
                       paste("Normalized Enrichment Score per",
                             switch(Z,
                                    lig="ligand",
                                    ct="cell line",
                                    ligct="ligand & cell line",
                                    rep="treatment condition")),
                       sep="\n"),
            main=temp_ds)
    if (length(CORS) > 300) {
      points(sapply(CORS,median),pch=".",col=alpha(temp_border,1))
    }
    abline(h=0,col="red",lty=2)
    if (temp_bot > 1.5) {
      if (Z %in% c("ligct","rep")) {
        mtext(names(temp_order),side=1,
              at=cumsum(sapply(temp_order,length)) - sapply(temp_order,length) / 2,
              las=2,xpd=NA,line=0.1,
              col=alpha(temp_border,1)[cumsum(sapply(temp_order,length))])
      } else {
        mtext(names(CORS),side=1,at=seq_along(CORS),
              las=2,xpd=NA,line=0.1,col=alpha(temp_border,1))
      }
    } else {
      mtext(switch(Z,
                   lig="Ligands",
                   ct="Cell lines",
                   ligct="Ligands per cell line",
                   rep="Treatment conditions"),
            side=1,line=0.2,cex=1.1,font=2)
    }
  }
}

```

Level 4 data (Z-scores) including inferred genes were used in GSEA analysis.  Pathway database was either GO-BP, Reactome, or All (`Human_GOBP_AllPathways_no_GO_iea_March_01_2021_entrezgene.gmt` from http://www.baderlab.org/GeneSets).  Inferred genes were included because without them, most pathways were not covered, which both artificially increased correlation of NES and drastically reduced power due to lack of genes.

Correlation of NES is a really dumb way of comparing enriched gene sets, but I already did the analysis, so here it is.


## GSEA correlation per cell line
```{r ct, fig.height=4,fig.width=9}
CORR_BXP("ct")
```


## GSEA correlation per ligand
```{r lig, fig.height=4,fig.width=9}
CORR_BXP("lig")
```


## GSEA correlation per ligand & cell line
```{r ligct, fig.height=4,fig.width=9}
CORR_BXP("ligct")
```


## GSEA correlation per treatment condition (replicate)
```{r rep, fig.height=4,fig.width=9}
CORR_BXP("rep")
```



## Comparison of all correlation distributions

As you can see, correlation of NES amongst replicates was the only thing that improved, which sort of makes sense as we don't expect low magnitude NES pathways to correlate at all.

```{r}
rm(list=ls())
```

```{r cors_per_lig, fig.height=4,fig.width=6}
source("~/Dropbox/GDB/line2user.R")

temp_GSEA <- unique(sapply(
  strsplit(list.files("~/Dropbox/GDB_archive/CMapCorr_files/",pattern="^corrGSEA_[A-Z]"),"_"),
  "[[",2))
for (GSEA in temp_GSEA) {
  
  temp <- load(paste0("~/Dropbox/GDB_archive/CMapCorr_files/corrGSEA_",GSEA,"_lig.RData"))
  cor_lig <- get(temp)
  temp <- load(paste0("~/Dropbox/GDB_archive/CMapCorr_files/corrGSEA_",GSEA,"_ligct.RData"))
  cor_ligct <- get(temp)
  temp <- load(paste0("~/Dropbox/GDB_archive/CMapCorr_files/corrGSEA_",GSEA,"_rep.RData"))
  cor_rep <- get(temp)
  rm(list=c(temp,"temp"))
  
  
  cor_ligct_lig <- sapply(strsplit(names(cor_ligct),"_"),"[[",1)
  cor_rep_lig <- sapply(strsplit(names(cor_rep),"_"),"[[",1)
  
  diff_ligct <- sapply(names(cor_lig),function(LIG) 
    mean(unlist(cor_ligct[cor_ligct_lig == LIG])) - mean(cor_lig[[LIG]]))
  mean_ligct <- sapply(names(cor_lig),function(LIG) 
    mean(c(unlist(cor_ligct[cor_ligct_lig == LIG]),cor_lig[[LIG]])))
  
  diff_rep <- sapply(names(cor_lig),function(LIG) 
    mean(unlist(cor_rep[cor_rep_lig == LIG])) - mean(cor_lig[[LIG]]))
  mean_rep <- sapply(names(cor_lig),function(LIG) 
    mean(c(unlist(cor_rep[cor_rep_lig == LIG]),cor_lig[[LIG]])))
  
  
  layout(rbind(1:2),widths=c(1,2))
  
  par(mar=c(5,3,1,0),mgp=2:0)
  boxplot(list(diff_ligct,diff_rep),
          col=c("red","dodgerblue"),xaxt="n",
          ylab="Difference in mean vs. by ligand")
  text(c(1,2),rep(line2user(0.5,1),2),
       c("By ligand / cell line","By replicate"),
       adj=1,srt=45,xpd=NA,cex=0.8)
  abline(h=0,lty=2)
  if (wilcox.test(diff_ligct)$p.value < 2.2e-16) {
    mtext("*",side=3,at=1,line=-0.3)
  }
  if (wilcox.test(diff_rep)$p.value < 2.2e-16) {
    mtext("*",side=3,at=2,line=-0.3)
  }
  
  par(mar=c(5,0,1,1))
  plot(NA,NA,xlab=NA,ylab=NA,yaxt="n",
       xlim=range(c(mean_ligct,mean_rep)),ylim=range(diff_ligct,diff_rep))
  mtext(paste("Mean pairwise Spearman correlation of",
              paste0("GSEA Normalized Enrichment Score (",GSEA,")"),
              sep="\n"),
        side=1,line=3.2)
  abline(h=0,lty=2)
  points(mean_ligct,diff_ligct,pch=20,col=alpha("red",0.5))
  points(mean_rep,diff_rep,pch=20,col=alpha("dodgerblue",0.5))
  legend("bottomright",c("By ligand / cell line","By replicate"),
         pch=20,col=alpha(c("red","dodgerblue"),0.5),cex=0.8,bty="n")
  
}
```

****

# Overlap of enriched gene sets

```{r}
rm(list=ls())
source("~/Dropbox/GDB/line2user.R")

load("~/Dropbox/GDB_archive/CMapCorr_files/lvl5_inputs_allgenes.RData")
temp_ligcountsperct <- sapply(unique(lvl5_data@cdesc$cell_id),function(CT)
  length(unique(lvl5_data@cdesc[lvl5_data@cdesc$cell_id == CT,"pert_iname"])))
ct9 <- names(temp_ligcountsperct)[temp_ligcountsperct > 100]
names(ct9) <- sapply(ct9,function(X) names(ct14)[ct14 == X])

temp_ctcountsperlig <- sapply(unique(lvl5_data@cdesc$pert_iname),function(LIG)
  length(unique(lvl5_data@cdesc[lvl5_data@cdesc$pert_iname == LIG &
                                  lvl5_data@cdesc$cell_id %in% ct9,"cell_id"])))
lig295 <- names(temp_ctcountsperlig)[temp_ctcountsperlig == 9]
lig295 <- sort(lig295)

temp_id <- rownames(lvl5_data@cdesc)[lvl5_data@cdesc$pert_iname %in% lig295 & 
                                       lvl5_data@cdesc$cell_id %in% ct9]
lvl5_data@mat <- lvl5_data@mat[,temp_id]
lvl5_data@cdesc <- lvl5_data@cdesc[temp_id,]
lvl5_data@cid <- temp_id
rm(list=c("lig16","ct14",grep("^temp",ls(),value=T)))
```


```{r}
load("~/Dropbox/GDB_archive/CMapCorr_files/povlpGSEA_ALL.RData")

Roverlap_ct <-sapply(Noverlap_ct,function(X)
  sapply(colnames(X),function(CTname)
    X[,CTname] / sum(lvl5_data@cdesc$cell_id == ct9[CTname])
  ),simplify=F)

Roverlap_lig <-sapply(Noverlap_lig,function(X)
  sapply(colnames(X),function(LIG)
    X[,LIG] / sum(lvl5_data@cdesc$pert_iname == LIG)
  ),simplify=F)

Roverlap_ligct <-sapply(Noverlap_ligct,function(X)
  sapply(colnames(X),function(LIGCT) {
    temp <- strsplit(LIGCT,"_")[[1]]
    X[,LIGCT] / sum(lvl5_data@cdesc$pert_iname == temp[1] &
                    lvl5_data@cdesc$cell_id == temp[2])
  }),simplify=F)


```

```{r eval=F,fig.width=9,fig.height=9}
cutFDR <- names(Noverlap_ct)

par(mfcol=c(3,3),mar=c(3,3,2,1),mgp=2:0)
for (FDR in cutFDR) {
  plot(as.vector(Roverlap_ct[[FDR]]),
       as.vector(-log10(Poverlap_ct[[FDR]])),
       xlim=c(0,1),pch=".")
  plot(as.vector(Roverlap_lig[[FDR]]),
       as.vector(-log10(Poverlap_lig[[FDR]])),
       xlim=c(0,1),pch=".")
  plot(as.vector(Roverlap_ligct[[FDR]]),
       as.vector(-log10(Poverlap_ligct[[FDR]])),
       xlim=c(0,1),pch=".")
}

```


> This is per gene set, which is a weird way of thinking about it.

```{r fig.width=9,fig.height=6}
cutFDR <- names(Noverlap_ct)
temp_p <- c(1,.5,.1,.05,.01,.001,.0001)

par(mfrow=c(3,2),mar=c(3,3,2,1),mgp=2:0)
for (FDR in cutFDR) {
  boxplot(list(CT=-log10(Poverlap_ct[[FDR]]),
               Lig=-log10(Poverlap_lig[[FDR]]),
               LigCT=-log10(Poverlap_ligct[[FDR]])),
          pch=".",yaxt="n",main=paste0("FDR: 0.",FDR),
          ylab="P significant gene sets by chance")
  axis(2,at=-log10(temp_p),labels=temp_p)
  
  
  boxplot(list(CT=Roverlap_ct[[FDR]],
               Lig=Roverlap_lig[[FDR]],
               LigCT=Roverlap_ligct[[FDR]]),
          pch=".",ylab="Proportion of overlapping")
}

```

# Mean NES

```{r eval=F}
rm(list=ls())
source("~/Dropbox/GDB/line2user.R")

load("~/Dropbox/GDB_archive/CMapCorr_files/lvl4_inputs_allgenes.RData")
lvl4_data <- lvl4_data_all; rm(lvl4_data_all)
temp_ligcountsperct <- sapply(unique(lvl4_data@cdesc$cell_id),function(CT)
  length(unique(lvl4_data@cdesc[lvl4_data@cdesc$cell_id == CT,"pert_iname"])))
ct9 <- names(temp_ligcountsperct)[temp_ligcountsperct > 100]
names(ct9) <- sapply(ct9,function(X) names(ct14)[ct14 == X])

temp_ctcountsperlig <- sapply(unique(lvl4_data@cdesc$pert_iname),function(LIG)
  length(unique(lvl4_data@cdesc[lvl4_data@cdesc$pert_iname == LIG &
                                  lvl4_data@cdesc$cell_id %in% ct9,"cell_id"])))
lig295 <- names(temp_ctcountsperlig)[temp_ctcountsperlig == 9]
lig295 <- sort(lig295)

temp_id <- rownames(lvl4_data@cdesc)[lvl4_data@cdesc$pert_iname %in% lig295 & 
                                       lvl4_data@cdesc$cell_id %in% ct9]
lvl4_data@mat <- lvl4_data@mat[,temp_id]
lvl4_data@cdesc <- lvl4_data@cdesc[temp_id,]
lvl4_data@cid <- temp_id
rm(list=c("lig16","ct14",grep("^temp",ls(),value=T)))
```

```{r fig.height=3,fig.width=6}
load("~/Dropbox/GDB_archive/CMapCorr_files/GSEA_lvl4_GOBP.RData")

par(mar=c(3,3,1,1),mgp=2:0)
hist(sapply(gsea_nes_GOBP[-1],"[[","NES"),breaks=50,main=NA,xlab="GSEA NES")
rm(gsea_nes_GOBP)
```


```{r}
GSEA <- "GOBP"
# load(paste0("~/Dropbox/GDB_archive/CMapCorr_files/pnesGSEA_",GSEA,".RData"))

load(paste0("~/Dropbox/GDB_archive/CMapCorr_files/pnesGSEA_GOBP_backup.RData"))
```


```{r fig.height=3,fig.width=9}
FDRcut <- 0.0001

tempMAT <- Pscore_CT[apply(Pscore_CT,1,function(X) any(abs(X) >= -log10(FDRcut))),]
temp_lim <- max(abs(tempMAT))
temp_hSample <- hclust(dist(t(tempMAT)),method="ward.D2") 
temp_hSet <- hclust(dist(tempMAT),method="ward.D2")
tempMAT <- tempMAT[temp_hSet$order,temp_hSample$order]

par(mar=c(2,6,0.5,0.5))
image(z=tempMAT,x=1:nrow(tempMAT),y=1:ncol(tempMAT),
      col=diverging_hcl(1000,palette="Blue-Red3"),
      breaks=seq(-temp_lim,temp_lim,length.out=1001),
      xaxt="n",yaxt="n",xlab=NA,ylab=NA)
mtext(colnames(tempMAT),side=2,line=0.1,at=1:ncol(tempMAT),las=2,cex=0.7)
mtext(paste(GSEA,"gene set enrichment"),side=1)
rm(list=grep("^temp",ls(),value=T))
```

```{r fig.height=6,fig.width=9}
tempMAT <- Pscore_Lig[,apply(Pscore_Lig,2,function(X) sum(abs(X) >= -log10(FDRcut)) >= 3)]
tempMAT <- tempMAT[apply(tempMAT,1,function(X) any(abs(X) >= -log10(FDRcut))),]
temp_lim <- max(abs(tempMAT))
temp_hSample <- hclust(dist(t(tempMAT)),method="ward.D2") 
temp_hSet <- hclust(dist(tempMAT),method="ward.D2")
tempMAT <- tempMAT[temp_hSet$order,temp_hSample$order]

par(mar=c(1.5,1.5,0.5,0.5))
image(z=tempMAT,x=1:nrow(tempMAT),y=1:ncol(tempMAT),
      col=diverging_hcl(1000,palette="Blue-Red3"),
      breaks=seq(-temp_lim,temp_lim,length.out=1001),
      xaxt="n",yaxt="n",xlab=NA,ylab=NA)
# mtext(colnames(tempMAT),side=2,line=0.1,at=1:ncol(tempMAT),las=2,cex=0.7)
mtext("Ligands",side=2,line=0.1)
mtext(paste(GSEA,"gene set enrichment"),side=1)
rm(list=grep("^temp",ls(),value=T))
```
```{r fig.height=9,fig.width=9}
tempMAT <- Pscore_LigCT[,apply(Pscore_LigCT,2,function(X) sum(abs(X) >= -log10(FDRcut)) >= 3)]
tempMAT <- tempMAT[apply(tempMAT,1,function(X) any(abs(X) >= -log10(FDRcut))),]
temp_lim <- max(abs(tempMAT))
temp_hSample <- hclust(dist(t(tempMAT)),method="ward.D2") 
temp_hSet <- hclust(dist(tempMAT),method="ward.D2")
tempMAT <- tempMAT[temp_hSet$order,temp_hSample$order]

par(mar=c(1.5,1.5,0.5,0.5))
image(z=tempMAT,x=1:nrow(tempMAT),y=1:ncol(tempMAT),
      col=diverging_hcl(1000,palette="Blue-Red3"),
      breaks=seq(-temp_lim,temp_lim,length.out=1001),
      xaxt="n",yaxt="n",xlab=NA,ylab=NA)
# mtext(colnames(tempMAT),side=2,line=0.1,at=1:ncol(tempMAT),las=2,cex=0.7)
mtext("Ligands per cell line",side=2,line=0.1)
mtext(paste(GSEA,"gene set enrichment"),side=1)
rm(list=grep("^temp",ls(),value=T))
```

```{r fig.height=3,fig.width=9}
par(mar=c(3,3,1,1),mgp=2:0)

boxplot(abs(Pscore_CT)[,order(colMeans(abs(Pscore_CT)))],
        pch=".",las=3,ylab="Mean NES |Significance|")
boxplot(abs(Pscore_Lig)[,order(colMeans(abs(Pscore_Lig)))],
        pch=".",las=3,xaxt="n",outline=F,
        xlab="Ligands",ylab="Mean NES |Significance|")
boxplot(abs(Pscore_LigCT)[,order(colMeans(abs(Pscore_LigCT)))],
        pch=".",las=3,xaxt="n",outline=F,
        xlab="Ligands per cell line",
        ylab="Mean NES |Significance|")


par(mfrow=c(1,2))
boxplot(list(CT=abs(Pscore_CT),
             Lig=abs(Pscore_Lig),
             LigCT=abs(Pscore_LigCT)),
        pch=".",horizontal=T,
        xlab="Mean NES |significance| (all)")
boxplot(list(CT=colMeans(abs(Pscore_CT)),
             Lig=colMeans(abs(Pscore_Lig)),
             LigCT=colMeans(abs(Pscore_LigCT))),
        pch=".",horizontal=T,
        xlab="Mean NES |significance| (means)")



```

# Mean signed FDR


```{r}
rm(list=ls())
source("~/Dropbox/GDB/line2user.R")

load("~/Dropbox/GDB_archive/CMapCorr_files/lvl4_inputs_allgenes.RData")
lvl4_data <- lvl4_data_all; rm(lvl4_data_all)
temp_ligcountsperct <- sapply(unique(lvl4_data@cdesc$cell_id),function(CT)
  length(unique(lvl4_data@cdesc[lvl4_data@cdesc$cell_id == CT,"pert_iname"])))
ct9 <- names(temp_ligcountsperct)[temp_ligcountsperct > 100]
names(ct9) <- sapply(ct9,function(X) names(ct14)[ct14 == X])

temp_ctcountsperlig <- sapply(unique(lvl4_data@cdesc$pert_iname),function(LIG)
  length(unique(lvl4_data@cdesc[lvl4_data@cdesc$pert_iname == LIG &
                                  lvl4_data@cdesc$cell_id %in% ct9,"cell_id"])))
lig295 <- names(temp_ctcountsperlig)[temp_ctcountsperlig == 9]
lig295 <- sort(lig295)

CID <- rownames(lvl4_data@cdesc)[lvl4_data@cdesc$pert_iname %in% lig295 & 
                                       lvl4_data@cdesc$cell_id %in% ct9]
CDESC <- lvl4_data@cdesc[CID,]

rm(list=c("lvl4_data","lig16","ct14",grep("^temp",ls(),value=T)))
```

Averaging "significance score" of each gene set across samples.  Score is `-log10(FDR) * sign(NES)`.  Example distribution:  

```{r fig.height=3,fig.width=6}
GSEA <- "GOBP"
load(paste0("~/Dropbox/GDB_archive/CMapCorr_files/pfdrGSEA_",GSEA,".RData"))

par(mar=c(4,3,1,1),mgp=2:0)
hist(GSEA_score,breaks=50,main=NA,xlab=NA)
mtext(paste("GSEA signed significance score",
                "-log10(FDR) * sign(NES)",sep="\n"),
      side=1,line=3)
```

The following heatmaps are showing gene set enrichment per sample group (cell lines, ligands, or ligands per cell line).  Values are signed FDR-corrected p-values representing the likelihood of seeing as extreme a value by chance when averaging signed GSEA FDR values across the sample group.  

## Cell line  

```{r fig.height=3,fig.width=9}
temp_bad <- which(abs(Mscore_CT) == max(abs(Mscore_CT)[Qscore_CT > 0.9]),arr.ind=T)
temp_good <- which(abs(Mscore_CT) == max(abs(Mscore_CT)[Qscore_CT <= min(Qscore_CT)]),arr.ind=T)

temp_bxp <- list(GSEA_score[temp_bad[1],CDESC$cell_id != ct9[temp_bad[2]]],
                 GSEA_score[temp_bad[1],CDESC$cell_id == ct9[temp_bad[2]]],
                 NULL,
                 GSEA_score[temp_good[1],CDESC$cell_id != ct9[temp_good[2]]],
                 GSEA_score[temp_good[1],CDESC$cell_id == ct9[temp_good[2]]])

layout(rbind(1:2),widths=c(3,2))
par(mar=c(3,4,1,1),mgp=2:0)
boxplot(temp_bxp,horizontal=T,pch=".",xlab="Signed GSEA FDR",yaxt="n",
        col=c(rep("red",2),NA,rep("dodgerblue",2)))
mtext(rep(c("All","Samples"),2),side=2,las=2,line=0.1,at=c(1,2,4,5))
points(sapply(temp_bxp[c(2,5)],mean),c(2,5),pch=20,col="yellow")

par(mar=c(3,3,1,1),mgp=2:0)
plot(Mscore_CT,-log10(Qscore_CT),pch=".",
     xlab="Mean signed GSEA FDR",ylab="-log10(FDR-corrected p)")
points(Mscore_CT[temp_bad],-log10(Qscore_CT)[temp_bad],
       pch=20,col="red")
points(Mscore_CT[temp_good],-log10(Qscore_CT)[temp_good],
       pch=20,col="dodgerblue")

rm(list=grep("^temp",ls(),value=T))
```

```{r fig.height=3,fig.width=9}
FDRcut <- 0.01

tempMAT <- -log10(Qscore_CT) * sign(Mscore_CT)
tempMAT <- tempMAT[apply(tempMAT,1,function(X) any(abs(X) >= -log10(FDRcut))),]
temp_lim <- max(abs(tempMAT))
temp_hSample <- hclust(dist(t(tempMAT)),method="ward.D2") 
temp_hSet <- hclust(dist(tempMAT),method="ward.D2")
tempMAT <- tempMAT[temp_hSet$order,temp_hSample$order]

par(mar=c(2,6,0.5,0.5))
image(z=tempMAT,x=1:nrow(tempMAT),y=1:ncol(tempMAT),
      col=diverging_hcl(1000,palette="Blue-Red3"),
      breaks=seq(-temp_lim,temp_lim,length.out=1001),
      xaxt="n",yaxt="n",xlab=NA,ylab=NA)
mtext(colnames(tempMAT),side=2,line=0.1,at=1:ncol(tempMAT),las=2,cex=0.7)
mtext(paste(GSEA,"gene set enrichment"),side=1)
mtext(paste0("|signed FDR| \u2264 ",FDRcut * 100,"%"),side=1,line=1)

rm(list=grep("^temp",ls(),value=T))
```

## Ligand  

```{r fig.height=3,fig.width=9}
temp_bad <- which(abs(Mscore_LIG) == max(abs(Mscore_LIG)[Qscore_LIG > 0.9]),arr.ind=T)
temp_good <- which(abs(Mscore_LIG) == max(abs(Mscore_LIG)[Qscore_LIG <= min(Qscore_LIG)]),arr.ind=T)

temp_bxp <- list(GSEA_score[temp_bad[1],CDESC$pert_iname != lig295[temp_bad[2]]],
                 GSEA_score[temp_bad[1],CDESC$pert_iname == lig295[temp_bad[2]]],
                 NULL,
                 GSEA_score[temp_good[1],CDESC$pert_iname != lig295[temp_good[2]]],
                 GSEA_score[temp_good[1],CDESC$pert_iname == lig295[temp_good[2]]])

layout(rbind(1:2),widths=c(3,2))
par(mar=c(3,4,1,1),mgp=2:0)
boxplot(temp_bxp,horizontal=T,pch=".",xlab="Signed GSEA FDR",yaxt="n",
        col=c(rep("red",2),NA,rep("dodgerblue",2)))
mtext(rep(c("All","Samples"),2),side=2,las=2,line=0.1,at=c(1,2,4,5))
points(sapply(temp_bxp[c(2,5)],mean),c(2,5),pch=20,col="yellow")

par(mar=c(3,3,1,1),mgp=2:0)
plot(Mscore_LIG,-log10(Qscore_LIG),pch=".",
     xlab="Mean signed GSEA FDR",ylab="-log10(FDR-corrected p)")
points(Mscore_LIG[temp_bad],-log10(Qscore_LIG)[temp_bad],
       pch=20,col="red")
points(Mscore_LIG[temp_good],-log10(Qscore_LIG)[temp_good],
       pch=20,col="dodgerblue")

# rm(list=grep("^temp",ls(),value=T))
```

```{r fig.height=6,fig.width=9}
tempMAT <- -log10(Qscore_LIG) * sign(Mscore_LIG)
tempMAT <- tempMAT[,apply(tempMAT,2,function(X) any(abs(X) >= -log10(FDRcut)))]
tempMAT <- tempMAT[apply(tempMAT,1,function(X) any(abs(X) >= -log10(FDRcut))),]
temp_lim <- max(abs(tempMAT))
temp_hSample <- hclust(dist(t(tempMAT)),method="ward.D2") 
temp_hSet <- hclust(dist(tempMAT),method="ward.D2")
tempMAT <- tempMAT[temp_hSet$order,temp_hSample$order]

par(mar=c(2,1.5,0.5,0.5))
image(z=tempMAT,x=1:nrow(tempMAT),y=1:ncol(tempMAT),
      col=diverging_hcl(1000,palette="Blue-Red3"),
      breaks=seq(-temp_lim,temp_lim,length.out=1001),
      xaxt="n",yaxt="n",xlab=NA,ylab=NA)
# mtext(colnames(tempMAT),side=2,line=0.1,at=1:ncol(tempMAT),las=2,cex=0.7)
mtext(paste0("Ligands (",ncol(tempMAT),"/",ncol(Mscore_LIG),")"),side=2,line=0.1)
mtext(paste(GSEA,"gene set enrichment"),side=1)
mtext(paste0("|signed FDR| \u2264 ",FDRcut * 100,"%"),side=1,line=1)

rm(list=grep("^temp",ls(),value=T))
```

## Ligand per cell line  

```{r fig.height=3,fig.width=9}
temp_bad <- which(abs(Mscore_LIGCT) == max(abs(Mscore_LIGCT)[Qscore_LIGCT > 0.9]),arr.ind=T)
temp_good <- which(abs(Mscore_LIGCT) == 
                     max(abs(Mscore_LIGCT)[Qscore_LIGCT <= min(Qscore_LIGCT)]),arr.ind=T)
try({
  temp_bad <- temp_bad[1,,drop=F]
})
try({
  temp_good <- temp_good[1,,drop=F]
})
temp_bad_cl <- strsplit(colnames(Mscore_LIGCT)[temp_bad[2]],"_")[[1]]
temp_good_cl <- strsplit(colnames(Mscore_LIGCT)[temp_good[2]],"_")[[1]]

temp_bxp <- list(GSEA_score[temp_bad[1],
                            CDESC$cell_id != temp_bad_cl[1] &
                              CDESC$pert_iname != temp_bad_cl[2]],
                 GSEA_score[temp_bad[1],
                            CDESC$cell_id == temp_bad_cl[1] &
                              CDESC$pert_iname == temp_bad_cl[2]],
                 NULL,
                 GSEA_score[temp_good[1],
                            CDESC$cell_id != temp_good_cl[1] &
                              CDESC$pert_iname != temp_good_cl[2]],
                 GSEA_score[temp_good[1],
                            CDESC$cell_id == temp_good_cl[1] &
                              CDESC$pert_iname == temp_good_cl[2]])

layout(rbind(1:2),widths=c(3,2))
par(mar=c(3,4,1,1),mgp=2:0)
boxplot(temp_bxp,horizontal=T,pch=".",xlab="Signed GSEA FDR",yaxt="n",
        col=c(rep("red",2),NA,rep("dodgerblue",2)))
mtext(rep(c("All","Samples"),2),side=2,las=2,line=0.1,at=c(1,2,4,5))
points(sapply(temp_bxp[c(2,5)],mean),c(2,5),pch=20,col="yellow")

par(mar=c(3,3,1,1),mgp=2:0)
plot(Mscore_LIGCT,-log10(Qscore_LIGCT),pch=".",
     xlab="Mean signed GSEA FDR",ylab="-log10(FDR-corrected p)")
points(Mscore_LIGCT[temp_bad],-log10(Qscore_LIGCT)[temp_bad],
       pch=20,col="red")
points(Mscore_LIGCT[temp_good],-log10(Qscore_LIGCT)[temp_good],
       pch=20,col="dodgerblue")

# rm(list=grep("^temp",ls(),value=T))
```

```{r fig.height=6,fig.width=9}
tempMAT <- -log10(Qscore_LIGCT) * sign(Mscore_LIGCT)
tempMAT <- tempMAT[,apply(tempMAT,2,function(X) any(abs(X) >= -log10(FDRcut)))]
tempMAT <- tempMAT[apply(tempMAT,1,function(X) any(abs(X) >= -log10(FDRcut))),]
temp_lim <- max(abs(tempMAT))
temp_hSample <- hclust(dist(t(tempMAT)),method="ward.D2") 
temp_hSet <- hclust(dist(tempMAT),method="ward.D2")
tempMAT <- tempMAT[temp_hSet$order,temp_hSample$order]

par(mar=c(2,1.5,0.5,0.5))
image(z=tempMAT,x=1:nrow(tempMAT),y=1:ncol(tempMAT),
      col=diverging_hcl(1000,palette="Blue-Red3"),
      breaks=seq(-temp_lim,temp_lim,length.out=1001),
      xaxt="n",yaxt="n",xlab=NA,ylab=NA)
# mtext(colnames(tempMAT),side=2,line=0.1,at=1:ncol(tempMAT),las=2,cex=0.7)
mtext(paste0("Ligands per cell line (",
             length(unique(sapply(strsplit(colnames(tempMAT),"_"),"[[",2))),
             "/295 unique ligands"),
      side=2,line=0.1)
mtext(paste(GSEA,"gene set enrichment"),side=1)
mtext(paste0("|signed FDR| \u2264 ",FDRcut * 100,"%"),side=1,line=1)

rm(list=grep("^temp",ls(),value=T))
```

```{r fig.height=3,fig.width=9}
par(mar=c(3,3,1,1),mgp=2:0)

boxplot(-log10(Qscore_CT)[,order(colMeans(-log10(Qscore_CT)))],
        pch=".",las=3,ylab="Mean FDR |Significance|")
boxplot(-log10(Qscore_LIG)[,order(colMeans(-log10(Qscore_LIG)))],
        pch=".",las=3,xaxt="n",outline=F,
        xlab="Ligands",ylab="Mean FDR |Significance|")
boxplot(-log10(Qscore_LIGCT)[,order(colMeans(-log10(Qscore_LIGCT)))],
        pch=".",las=3,xaxt="n",outline=F,
        xlab="Ligands per cell line",
        ylab="Mean FDR |Significance|")


par(mfrow=c(1,2))
boxplot(list(CT=-log10(Qscore_CT),
             Lig=-log10(Qscore_LIG),
             LigCT=-log10(Qscore_LIGCT)),
        pch=".",horizontal=T,
        xlab="Mean FDR |significance| (all)")
boxplot(list(CT=colMeans(-log10(Qscore_CT)),
             Lig=colMeans(-log10(Qscore_LIG)),
             LigCT=colMeans(-log10(Qscore_LIGCT))),
        pch=".",horizontal=T,
        xlab="Mean FDR |significance| (means)")



```
```{r fig.height=5,fig.width=9}
allCuts <- seq(4,0,by=-0.01)

# By sample ----

Nhits_CT <- sapply(allCuts,function(CUT) colMeans(-log10(Qscore_CT) >= CUT))
Nhits_Lig <- sapply(allCuts,function(CUT) colMeans(-log10(Qscore_LIG) >= CUT))
Nhits_LigCT <- sapply(allCuts,function(CUT) colMeans(-log10(Qscore_LIGCT) >= CUT))

# par(mfrow=c(3,1),mar=c(3,3,1,1),mgp=2:0)
# boxplot(Nhits_CT,pch=".")
# boxplot(Nhits_Lig,pch=".")
# boxplot(Nhits_LigCT,pch=".")


par(mfrow=c(1,2),mar=c(3,3,1,1),mgp=2:0)
plot(NA,NA,xlim=c(4,0),ylim=c(0,1),xaxt="n",
     ylab="Proportion of sample set |mean signed FDR| \u2264 x",xlab=NA)
axis(side=1,labels=c(.0001,.001,.01,.05,.1,.5,1),
     at=-log10(c(.0001,.001,.01,.05,.1,.5,1)))
mtext("Gene set |mean signed FDR| threshold",side=1,line=2)

# CT ----
lines(allCuts,colMeans(Nhits_CT),lty=1,lwd=2,
      col=qualitative_hcl(3,palette="dark3")[1])
temp <- apply(apply(Nhits_CT,2,fivenum)[c(2,4),],1,function(X)
  lines(allCuts,X,lty=3,lwd=2,col=qualitative_hcl(3,palette="dark3")[1]))

# Lig ----
lines(allCuts,colMeans(Nhits_Lig),lty=1,lwd=2,
      col=qualitative_hcl(3,palette="dark3")[2])
temp <- apply(apply(Nhits_Lig,2,fivenum)[c(2,4),],1,function(X)
  lines(allCuts,X,lty=3,lwd=2,col=qualitative_hcl(3,palette="dark3")[2]))

# LigCT ----
lines(allCuts,colMeans(Nhits_LigCT),lty=1,lwd=2,
      col=qualitative_hcl(3,palette="dark3")[3])
temp <- apply(apply(Nhits_LigCT,2,fivenum)[c(2,4),],1,function(X)
  lines(allCuts,X,lty=3,lwd=2,col=qualitative_hcl(3,palette="dark3")[3]))

legend("topleft",legend=c("Cell line","Ligand","Ligand per cell line","Mean","Quartiles"),
       bty="n",col=c(qualitative_hcl(3,palette="dark3"),1,1),lty=c(rep(1,4),3),lwd=2)



# All ----

Nhits_CT <- sapply(allCuts,function(CUT) mean(-log10(Qscore_CT) >= CUT))
Nhits_Lig <- sapply(allCuts,function(CUT) mean(-log10(Qscore_LIG) >= CUT))
Nhits_LigCT <- sapply(allCuts,function(CUT) mean(-log10(Qscore_LIGCT) >= CUT))

plot(NA,NA,xlim=c(4,0),ylim=c(0,1),xaxt="n",
     ylab="Proportion of all |mean signed FDR| \u2264 x",xlab=NA)
axis(side=1,labels=c(.0001,.001,.01,.05,.1,.5,1),
     at=-log10(c(.0001,.001,.01,.05,.1,.5,1)))
mtext("Gene set |mean signed FDR| threshold",side=1,line=2)
lines(allCuts,Nhits_CT,lty=1,lwd=2,
      col=qualitative_hcl(3,palette="dark3")[1])
lines(allCuts,Nhits_Lig,lty=1,lwd=2,
      col=qualitative_hcl(3,palette="dark3")[2])
lines(allCuts,Nhits_LigCT,lty=1,lwd=2,
      col=qualitative_hcl(3,palette="dark3")[3])
legend("topleft",legend=c("Cell line","Ligand","Ligand per cell line"),
       bty="n",col=qualitative_hcl(3,palette="dark3"),lty=1,lwd=2)


```

P-value for difference in CDFs between cell line and ligand by KS-test: `ks.test(as.vector(Qscore_LIG),as.vector(Qscore_CT))$p.value`  

P-value for difference in CDFs between ligand and ligand per cell line by KS-test: `ks.test(as.vector(Qscore_LIG),as.vector(Qscore_LIGCT))$p.value`  


```{r}
rm(list=c("FDRcut",grep("^Pscore",ls(),value=T)))
```

## Compare mean signed FDR values  
```{r, eval=F}
load("~/Dropbox/GDB_archive/CMapCorr_files/phitsGSEA_GOBP.RData")
```

```{r, eval=F}
plot(Nhits_LIGCT,-log10(Phits_LIGCT),pch=".",cex=2,
     col=sapply(sequential_hcl(nrow(Nhits_LIGCT),palette="Batlow",alpha=0.5),
                rep,times=ncol(Nhits_LIGCT)))
```

```{r eval=F, fig.width=9,fig.height=4}
par(mfrow=c(3,1),mar=c(3,3,1,1),mgp=2:0)
hist(-log10(Phits_CT),breaks=50)
abline(v=mean(-log10(Phits_CT)),col="dodgerblue")
hist(-log10(Phits_LIG),breaks=50)
abline(v=mean(-log10(Phits_LIG)),col="dodgerblue")
hist(-log10(Phits_LIGCT),breaks=50)
abline(v=mean(-log10(Phits_LIGCT)),col="dodgerblue")

```



