---
title: "Correlation between Z-scores"
---
  
```{r setup, echo=F,message=F,warning=F}
library(scales)
library(colorspace)
source("~/Dropbox/GDB/line2user.R")
.PAR <- par(no.readonly=T)
knitr::opts_chunk$set(
  echo=F,
  message=F,
  warning=F
)
```

# Supplemental Figure

```{r}
temp <- load("~/Dropbox/GDB_archive/CMapCorr_files/corr_lvl3_lig295.RData") 
corr_lvl3 <- list(CT=unlist(lvl3_ct_corr,use.names=F),
                  LIG=unlist(lvl3_lig_corr,use.names=F),
                  LIGCT=unlist(lvl3_lig_corr_ct,use.names=F),
                  REP=unlist(lvl3_lig_corr_tx,use.names=F))
rm(list=c(temp,"temp"))

corr_lvl4 <- sapply(list.files("~/Dropbox/GDB_archive/CMapCorr_files","^lig295_corr_[^a]",full.names=T),
                    function(X) {
                      temp <- load(X)
                      return(unlist(get(temp),use.names=F))
                    },simplify=F)
names(corr_lvl4) <- names(corr_lvl3)

```

```{r echo=TRUE}
signif(sapply(corr_lvl3,mean),3)
signif(sapply(corr_lvl3,median),3)
signif(sapply(corr_lvl4,mean),3)
signif(sapply(corr_lvl4,median),3)
```


```{r suppfigAB, fig.height=4,fig.width=8}
par(mfrow=c(1,2),mar=c(4.5,1.5,1,0.5),mgp=2:0)

temp_dens <- sapply(corr_lvl3,density,simplify=F)
# temp_dens <- rev(sapply(corr_lvl3,density,simplify=F))
temp_zero <- max(sapply(temp_dens,"[[","y")) * seq(0,0.5,length.out=length(temp_dens))


plot(NA,NA,xlim=range(sapply(temp_dens,"[[","x")),
     ylim=c(0,max(sapply(temp_dens,"[[","y")) + max(temp_zero)),
     xlab=NA,ylab=NA,yaxt="n",main=NA)
abline(h=temp_zero)
mtext("Density",side=2,line=0.1)
mtext(paste("Pairwise Spearman correlation between",
            "quantile-normalized transcript abundances",sep="\n"),
      side=1,line=3.1)
temp <- sapply(rev(seq_along(temp_dens)),function(X)
  polygon(temp_dens[[X]]$x,temp_dens[[X]]$y + temp_zero[X],lwd=2,
          border=qualitative_hcl(length(temp_dens),palette="dark3")[X],
          col=qualitative_hcl(length(temp_dens),palette="dark3",alpha=0.7)[X]))
text(line2user(-0.1,2),temp_zero,adj=c(0,-0.3),
     labels=c(CT="Cell line",LIG="Ligand",
              LIGCT="Ligand per cell line",REP="Replicate")[names(temp_dens)],
     col=qualitative_hcl(length(temp_dens),palette="dark3"))  
mtext("A",side=3,line=-0.5,at=line2user(0.75,2),font=2,cex=1.5)


temp_dens <- sapply(corr_lvl4,density,simplify=F)
# temp_dens <- rev(sapply(corr_lvl4,density,simplify=F))
temp_zero <- max(sapply(temp_dens,"[[","y")) * seq(0,0.5,length.out=length(temp_dens))

plot(NA,NA,xlim=range(sapply(temp_dens,"[[","x")),
     ylim=c(0,max(sapply(temp_dens,"[[","y")) + max(temp_zero)),
     xlab=NA,ylab=NA,yaxt="n",main=NA)
abline(h=temp_zero)
mtext("Density",side=2,line=0.1)
mtext(paste("Pairwise Spearman correlation between",
            "ligand response Z-scores",sep="\n"),
      side=1,line=3.1)
temp <- sapply(rev(seq_along(temp_dens)),function(X)
  polygon(temp_dens[[X]]$x,temp_dens[[X]]$y + temp_zero[X],lwd=2,
          border=qualitative_hcl(length(temp_dens),palette="dark3")[X],
          col=qualitative_hcl(length(temp_dens),palette="dark3",alpha=0.7)[X]))
# text(line2user(-0.1,2),temp_zero,adj=c(0,-0.3),
#      labels=c(CT="Cell line",LIG="Ligand",
#               LIGCT="Ligand per cell line",REP="Replicate")[names(temp_dens)],
#      col=qualitative_hcl(length(temp_dens),palette="dark3"))  
mtext("B",side=3,line=-0.5,at=line2user(0.75,2),font=2,cex=1.5)
```



```{r}
rm(list=ls())
source("~/Dropbox/GDB/line2user.R")
temp <- load("~/Dropbox/GDB_archive/CMapCorr_files/lig295_corr_lig.RData")
cor_lig <- get(temp)
temp <- load("~/Dropbox/GDB_archive/CMapCorr_files/lig295_corr_ligct.RData")
cor_ligct <- get(temp)
temp <- load("~/Dropbox/GDB_archive/CMapCorr_files/lig295_corr_rep.RData")
cor_rep <- get(temp)
rm(list=c(temp,"temp"))
```

```{r}
cor_ligct_lig <- sapply(strsplit(names(cor_ligct),"_"),"[[",1)
cor_rep_lig <- sapply(strsplit(names(cor_rep),"_"),"[[",1)

# boxplot(list(LIG=cor_lig[["AGT"]],
#              LIGCT=unlist(cor_ligct[cor_ligct_lig == "AGT"]),
#              REP=unlist(cor_rep[cor_rep_lig == "AGT"])))

diff_ligct <- sapply(names(cor_lig),function(LIG) 
  mean(unlist(cor_ligct[cor_ligct_lig == LIG])) - mean(cor_lig[[LIG]]))
mean_ligct <- sapply(names(cor_lig),function(LIG) 
  mean(c(unlist(cor_ligct[cor_ligct_lig == LIG]),cor_lig[[LIG]])))

diff_rep <- sapply(names(cor_lig),function(LIG) 
  mean(unlist(cor_rep[cor_rep_lig == LIG])) - mean(cor_lig[[LIG]]))
mean_rep <- sapply(names(cor_lig),function(LIG) 
  mean(c(unlist(cor_rep[cor_rep_lig == LIG]),cor_lig[[LIG]])))

temp_all <- sapply(names(cor_lig),function(X)
  list(Lig=cor_lig[[X]],
       LigCT=unlist(cor_ligct[cor_ligct_lig == X]),
       Rep=unlist(cor_rep[cor_rep_lig == X])),
  simplify=F)
temp_all_stat <- sapply(temp_all,function(X) 
  c(LigCT=wilcox.test(X$LigCT,X$Lig)$p.value,
       Rep=wilcox.test(X$Rep,X$Lig)$p.value))


# par(mar=c(5.5,6,2.5,1),mgp=2:0)
# plot(NA,NA,xlim=c(0,length(temp_labels) * 4),ylim=c(-1,1),
#      xlab=NA,ylab=NA,xaxt="n",yaxs="i")
# abline(h=0,lty=2,col="red")
# 
# boxplot(unlist(temp_bxp,recursive=F),xaxt="n",
#   col=as.vector(sapply(qualitative_hcl(length(temp_labels),palette="dark3"),rep,3)),
#   pch=20,add=T,at=seq(length(temp_labels) * 4)[-seq(4,by=4,length.out=length(temp_labels))])
# mtext(temp_labels,side=3,line=0.1,
#       col=qualitative_hcl(length(temp_labels),palette="dark3"),
#       at=seq(2,by=4,length.out=length(temp_labels)))
# 
# mtext("Pairwise Spearman correlation of Z-scores",side=2,line=2.2)
# text(seq(length(temp_labels) * 4)[-seq(4,by=4,length.out=length(temp_labels))],
#      line2user(0.5,1),
#      rep(c("By ligand","By ligand / cell line","By replicate"),length(temp_labels)),
#      xpd=NA,srt=40,adj=1)
# mtext("B",font=2,cex=1.5,line=0.5,side=3,at=line2user(4.5,2))
# mtext("*",side=3,line=-1.5,at=c(2,3,6,7,10,11)[as.vector(temp_stat) <= 0.05])
```

```{r suppfigC, fig.height=5,fig.width=8}
layout(rbind(1:2),widths=c(1,2))
par(mar=c(5,3.5,1,0),mgp=2:0)
boxplot(list(diff_ligct,diff_rep),col=c("red","dodgerblue"),xaxt="n",pch=".",cex=2)
abline(h=0,lty=2)
mtext("Difference in mean correlation vs. by ligand",side=2,line=2.2)
text(c(1,2),rep(line2user(0.5,1),2),
       c("By ligand / cell line","By replicate"),
       adj=1,srt=35,xpd=NA,cex=0.9)
if (wilcox.test(diff_ligct)$p.value < 2.2e-16) {
  mtext("*",side=3,at=1,font=2,line=-0.3)
}
if (wilcox.test(diff_rep)$p.value < 2.2e-16) {
  mtext("*",side=3,at=2,font=2,line=-0.3)
}
mtext("C",font=2,cex=1.5,line=-0.5,side=3,at=line2user(2.75,2))

par(mar=c(5,0,1,4))
plot(NA,NA,xlab=NA,ylab=NA,yaxt="n",
     xlim=range(c(mean_ligct,mean_rep)),ylim=range(diff_ligct,diff_rep))
mtext("Mean pairwise Spearman correlation of Z-scores",side=1,line=2.2)
axis(4)
mtext("Difference in mean correlation vs. by ligand",side=4,line=2.2)
abline(h=0,lty=2)
points(mean_rep[temp_all_stat["Rep",names(mean_rep)] > 0.05],
       diff_rep[temp_all_stat["Rep",names(mean_rep)] > 0.05],
       pch=20,col=alpha("dodgerblue",0.2))
points(mean_ligct[temp_all_stat["LigCT",names(mean_ligct)] > 0.05],
       diff_ligct[temp_all_stat["LigCT",names(mean_ligct)] > 0.05],
       pch=20,col=alpha("red",0.2))
points(mean_rep[temp_all_stat["Rep",names(mean_rep)] <= 0.05],
       diff_rep[temp_all_stat["Rep",names(mean_rep)] <= 0.05],
       pch=19,col=alpha("dodgerblue",0.5))
points(mean_ligct[temp_all_stat["LigCT",names(mean_ligct)] <= 0.05],
       diff_ligct[temp_all_stat["LigCT",names(mean_ligct)] <= 0.05],
       pch=19,col=alpha("red",0.5))
legend("bottomright",bty="n",
       c("By ligand / cell line","By replicate","\u0394 ligand p \u2264 0.05"),
       pch=c(20,20,19),col=c(alpha(c("red","dodgerblue","black"),0.5)))



```

"*" p < 2.2e-16 by Wilcoxon signed-rank test.  
Mean improvement (LIGCT): `r mean(diff_ligct)`  
Mean improvement (REP): `r mean(diff_rep)`  




****

# OLD


## Figure 1a

```{r fig1a, fig.height=4,fig.width=9}
Z <- "lig"
LIG <- "lig295"
DAT <- "lvl4_data"

temp <- switch(grepl("all",DAT) + 1,
               paste0(LIG,"_corr_",Z,".RData"),
               paste0(LIG,"_corr_allgenes_",Z,".RData"))
load(paste0("~/Dropbox/GDB_archive/CMapCorr_files/",temp))

if (Z %in% c("ligct","rep")) {
  temp_lig <- as.factor(sub("_[A-Za-z0-9_]+$","",names(CORS)))
  temp_ord1 <- tapply(CORS,temp_lig,function(X) names(X)[order(sapply(X,median))])
  temp_ord2 <- order(tapply(CORS,temp_lig,function(X) median(unlist(X))))
  temp_order <- temp_ord1[temp_ord2]
  CORS <- CORS[unlist(temp_order)]
  temp_border <- qualitative_hcl(length(temp_order),palette="dark3")
  temp_border <- unlist(mapply(rep,
                               x=temp_border,
                               times=sapply(temp_order,length),
                               SIMPLIFY=F))
  temp_col <- alpha(temp_border,0.5)
  temp_bot <- switch(1 + (length(temp_order) < 20),1.5,4)
} else {
  temp_order <- order(sapply(CORS,median))
  CORS <- CORS[temp_order]
  temp_border <- qualitative_hcl(length(CORS),palette="dark3")
  temp_col <- qualitative_hcl(length(CORS),palette="dark3",alpha=0.5)
  temp_bot <- switch(1 + (length(CORS) < 20),1.5,4)
}
if (length(CORS) > 300) {
  temp_border <- alpha(temp_border,0.5)
  temp_col <- alpha(temp_col,0.2)
}      
par(mar=c(temp_bot,5,2,1),mgp=2:0)
boxplot(CORS,outline=T,pch=".",xaxt="n",yaxs="i",ylim=c(-1,1),
        border=temp_border,col=temp_col,
        ylab=paste("Pairwise Spearman correlation of",
                   paste("Z-scores per",
                         switch(Z,
                                lig="ligand",
                                ct="cell line",
                                ligct="ligand & cell line",
                                rep="treatment condition")),
                   sep="\n"))
if (length(CORS) > 300) {
  points(sapply(CORS,median),pch=".",col=alpha(temp_border,1))
}
abline(h=0,col="red",lty=2)
if (temp_bot > 1.5) {
  if (Z %in% c("ligct","rep")) {
    mtext(names(temp_order),side=1,
          at=cumsum(sapply(temp_order,length)) - sapply(temp_order,length) / 2,
          las=2,xpd=NA,line=0.1,
          col=alpha(temp_border,1)[cumsum(sapply(temp_order,length))])
  } else {
    mtext(names(CORS),side=1,at=seq_along(CORS),
          las=2,xpd=NA,line=0.1,col=alpha(temp_border,1))
  }
} else {
  mtext(switch(Z,
               lig="Ligands",
               ct="Cell lines",
               ligct="Ligands per cell line",
               rep="Treatment conditions"),
        side=1,line=0.2,cex=1.1,font=2)
}
mtext("A",font=2,cex=1.5,line=0.5,side=3,at=line2user(4,2))
```

Mean of median pairwise Spearman correlation of Z-scores per ligand is `r mean(sapply(CORS,median))`.


## Figure 1b/c

```{r load_all_cors}
rm(list=ls())
source("~/Dropbox/GDB/line2user.R")
temp <- load("~/Dropbox/GDB_archive/CMapCorr_files/lig295_corr_lig.RData")
cor_lig <- get(temp)
temp <- load("~/Dropbox/GDB_archive/CMapCorr_files/lig295_corr_ligct.RData")
cor_ligct <- get(temp)
temp <- load("~/Dropbox/GDB_archive/CMapCorr_files/lig295_corr_rep.RData")
cor_rep <- get(temp)
rm(list=c(temp,"temp"))
```

```{r cors_per_lig, fig.height=4,fig.width=9}
cor_ligct_lig <- sapply(strsplit(names(cor_ligct),"_"),"[[",1)
cor_rep_lig <- sapply(strsplit(names(cor_rep),"_"),"[[",1)

# boxplot(list(LIG=cor_lig[["AGT"]],
#              LIGCT=unlist(cor_ligct[cor_ligct_lig == "AGT"]),
#              REP=unlist(cor_rep[cor_rep_lig == "AGT"])))

diff_ligct <- sapply(names(cor_lig),function(LIG) 
  mean(unlist(cor_ligct[cor_ligct_lig == LIG])) - mean(cor_lig[[LIG]]))
mean_ligct <- sapply(names(cor_lig),function(LIG) 
  mean(c(unlist(cor_ligct[cor_ligct_lig == LIG]),cor_lig[[LIG]])))

diff_rep <- sapply(names(cor_lig),function(LIG) 
  mean(unlist(cor_rep[cor_rep_lig == LIG])) - mean(cor_lig[[LIG]]))
mean_rep <- sapply(names(cor_lig),function(LIG) 
  mean(c(unlist(cor_rep[cor_rep_lig == LIG]),cor_lig[[LIG]])))

temp_labels <- c("ADM","IL1A","TNF")

layout(rbind(1:3),widths=c(2,3,1))
par(mar=c(5.5,6,2.5,1),mgp=2:0)
plot(NA,NA,xlim=c(0,length(temp_labels) * 4),ylim=c(-1,1),
     xlab=NA,ylab=NA,xaxt="n",yaxs="i")
abline(h=0,lty=2,col="red")
boxplot(unlist(sapply(temp_labels,function(X) 
  list(Lig=cor_lig[[X]],
       LigCT=unlist(cor_ligct[cor_ligct_lig == X]),
       Rep=unlist(cor_rep[cor_rep_lig == X])),
  simplify=F),recursive=F),xaxt="n",
  col=as.vector(sapply(qualitative_hcl(length(temp_labels),palette="dark3"),rep,3)),
  pch=20,add=T,at=seq(length(temp_labels) * 4)[-seq(4,by=4,length.out=length(temp_labels))])
mtext(temp_labels,side=3,line=0.1,
      col=qualitative_hcl(length(temp_labels),palette="dark3"),
      at=seq(2,by=4,length.out=length(temp_labels)))
mtext("Pairwise Spearman correlation of Z-scores",side=2,line=2.2)
text(seq(length(temp_labels) * 4)[-seq(4,by=4,length.out=length(temp_labels))],
     line2user(0.5,1),
     rep(c("By ligand","By ligand / cell line","By replicate"),length(temp_labels)),
     xpd=NA,srt=40,adj=1)
mtext("B",font=2,cex=1.5,line=0.5,side=3,at=line2user(4.5,2))


par(mar=c(4,5,2.5,0))
plot(NA,NA,xlab=NA,ylab=NA,
     xlim=range(c(mean_ligct,mean_rep)),ylim=range(diff_ligct,diff_rep))
mtext("Mean pairwise Spearman correlation of Z-scores",side=1,line=2.2)
mtext("Difference in mean vs. by ligand",side=2,line=2.2)
abline(h=0,lty=2)
points(mean_ligct,diff_ligct,pch=20,col=alpha("red",0.5))
text(mean_ligct[temp_labels],diff_ligct[temp_labels],
     temp_labels,col="red",pos=c(2,4,4),offset=0.2)
points(mean_rep,diff_rep,pch=20,col=alpha("dodgerblue",0.5))
text(mean_rep[temp_labels],diff_rep[temp_labels],
     temp_labels,col="dodgerblue",pos=c(2,4,4),offset=0.2)
legend("topright",c("By ligand / cell line","By replicate"),
       pch=20,col=alpha(c("red","dodgerblue"),0.5))
mtext("C",font=2,cex=1.5,line=0.5,side=3,at=line2user(4,2))


par(mar=c(4,0,2.5,4))
boxplot(list(diff_ligct,diff_rep),col=c("red","dodgerblue"),
        yaxt="n",xaxt="n")
abline(h=0,lty=2)
axis(4)
mtext("Difference in mean vs. by ligand",side=4,line=2.2)
if (wilcox.test(diff_ligct)$p.value < 2.2e-16) {
  mtext("*",side=3,at=1,line=-0.5)
}
if (wilcox.test(diff_rep)$p.value < 2.2e-16) {
  mtext("*",side=3,at=2,line=-0.5)
}

```

The * denotes p < 2.2e-16 from Wilcoxon signed-rank test vs. mu of 0.  
Mean improvement by ligand / cell line: `r mean(diff_ligct)`  
Mean improvement by replicate: `r mean(diff_rep)`  


## New version?

```{r load_all_cors2}
rm(list=ls())
source("~/Dropbox/GDB/line2user.R")
temp <- load("~/Dropbox/GDB_archive/CMapCorr_files/lig295_corr_lig.RData")
cor_lig <- get(temp)
temp <- load("~/Dropbox/GDB_archive/CMapCorr_files/lig295_corr_ligct.RData")
cor_ligct <- get(temp)
temp <- load("~/Dropbox/GDB_archive/CMapCorr_files/lig295_corr_rep.RData")
cor_rep <- get(temp)
rm(list=c(temp,"temp"))
```

```{r cors_per_lig2, fig.height=5,fig.width=8}
cor_ligct_lig <- sapply(strsplit(names(cor_ligct),"_"),"[[",1)
cor_rep_lig <- sapply(strsplit(names(cor_rep),"_"),"[[",1)

# boxplot(list(LIG=cor_lig[["AGT"]],
#              LIGCT=unlist(cor_ligct[cor_ligct_lig == "AGT"]),
#              REP=unlist(cor_rep[cor_rep_lig == "AGT"])))

diff_ligct <- sapply(names(cor_lig),function(LIG) 
  mean(unlist(cor_ligct[cor_ligct_lig == LIG])) - mean(cor_lig[[LIG]]))
mean_ligct <- sapply(names(cor_lig),function(LIG) 
  mean(c(unlist(cor_ligct[cor_ligct_lig == LIG]),cor_lig[[LIG]])))

diff_rep <- sapply(names(cor_lig),function(LIG) 
  mean(unlist(cor_rep[cor_rep_lig == LIG])) - mean(cor_lig[[LIG]]))
mean_rep <- sapply(names(cor_lig),function(LIG) 
  mean(c(unlist(cor_rep[cor_rep_lig == LIG]),cor_lig[[LIG]])))

temp_all <- sapply(names(cor_lig),function(X)
  list(Lig=cor_lig[[X]],
       LigCT=unlist(cor_ligct[cor_ligct_lig == X]),
       Rep=unlist(cor_rep[cor_rep_lig == X])),
  simplify=F)
temp_all_stat <- sapply(temp_all,function(X) 
  c(LigCT=wilcox.test(X$LigCT,X$Lig)$p.value,
       Rep=wilcox.test(X$Rep,X$Lig)$p.value))


# par(mar=c(5.5,6,2.5,1),mgp=2:0)
# plot(NA,NA,xlim=c(0,length(temp_labels) * 4),ylim=c(-1,1),
#      xlab=NA,ylab=NA,xaxt="n",yaxs="i")
# abline(h=0,lty=2,col="red")
# 
# boxplot(unlist(temp_bxp,recursive=F),xaxt="n",
#   col=as.vector(sapply(qualitative_hcl(length(temp_labels),palette="dark3"),rep,3)),
#   pch=20,add=T,at=seq(length(temp_labels) * 4)[-seq(4,by=4,length.out=length(temp_labels))])
# mtext(temp_labels,side=3,line=0.1,
#       col=qualitative_hcl(length(temp_labels),palette="dark3"),
#       at=seq(2,by=4,length.out=length(temp_labels)))
# 
# mtext("Pairwise Spearman correlation of Z-scores",side=2,line=2.2)
# text(seq(length(temp_labels) * 4)[-seq(4,by=4,length.out=length(temp_labels))],
#      line2user(0.5,1),
#      rep(c("By ligand","By ligand / cell line","By replicate"),length(temp_labels)),
#      xpd=NA,srt=40,adj=1)
# mtext("B",font=2,cex=1.5,line=0.5,side=3,at=line2user(4.5,2))
# mtext("*",side=3,line=-1.5,at=c(2,3,6,7,10,11)[as.vector(temp_stat) <= 0.05])


layout(rbind(1:2),widths=c(1,2))
par(mar=c(5,4,1,0),mgp=2:0)
boxplot(list(diff_ligct,diff_rep),col=c("red","dodgerblue"),xaxt="n",pch=".",cex=2)
abline(h=0,lty=2)
mtext("Difference in mean correlation vs. by ligand",side=2,line=2.2)
text(c(1,2),rep(line2user(0.5,1),2),
       c("By ligand / cell line","By replicate"),
       adj=1,srt=35,xpd=NA,cex=0.9)
if (wilcox.test(diff_ligct)$p.value < 2.2e-16) {
  mtext("*",side=3,at=1,font=2,line=-0.3)
}
if (wilcox.test(diff_rep)$p.value < 2.2e-16) {
  mtext("*",side=3,at=2,font=2,line=-0.3)
}
mtext("A",font=2,cex=1.5,line=-0.5,side=3,at=line2user(3.5,2))

par(mar=c(5,0,1,4))
plot(NA,NA,xlab=NA,ylab=NA,yaxt="n",
     xlim=range(c(mean_ligct,mean_rep)),ylim=range(diff_ligct,diff_rep))
mtext("Mean pairwise Spearman correlation of Z-scores",side=1,line=2.2)
axis(4)
mtext("Difference in mean correlation vs. by ligand",side=4,line=2.2)
abline(h=0,lty=2)
points(mean_rep[temp_all_stat["Rep",names(mean_rep)] > 0.05],
       diff_rep[temp_all_stat["Rep",names(mean_rep)] > 0.05],
       pch=20,col=alpha("dodgerblue",0.2))
points(mean_ligct[temp_all_stat["LigCT",names(mean_ligct)] > 0.05],
       diff_ligct[temp_all_stat["LigCT",names(mean_ligct)] > 0.05],
       pch=20,col=alpha("red",0.2))
points(mean_rep[temp_all_stat["Rep",names(mean_rep)] <= 0.05],
       diff_rep[temp_all_stat["Rep",names(mean_rep)] <= 0.05],
       pch=19,col=alpha("dodgerblue",0.5))
points(mean_ligct[temp_all_stat["LigCT",names(mean_ligct)] <= 0.05],
       diff_ligct[temp_all_stat["LigCT",names(mean_ligct)] <= 0.05],
       pch=19,col=alpha("red",0.5))
legend("bottomright",bty="n",
       c("By ligand / cell line","By replicate","\u0394 ligand p \u2264 0.05"),
       pch=c(20,20,19),col=c(alpha(c("red","dodgerblue","black"),0.5)))



```


"*" p < 2.2e-16 by Wilcoxon signed-rank test.  

