---
title: "Correlation between Z-scores"
---
  
```{r setup, echo=F,message=F,warning=F}
library(scales)
library(colorspace)
source("~/Dropbox/GDB/line2user.R")
.PAR <- par(no.readonly=T)
knitr::opts_chunk$set(
  echo=F,
  message=F,
  warning=F
)
```


```{r FX}
CORR_BXP <- function(Z) {
  for (LIG in c("lig16","lig295")) {
    for (DAT in c("lvl4_data","lvl4_data_all")) {
      temp <- switch(grepl("all",DAT) + 1,
                     paste0(LIG,"_corr_",Z,".RData"),
                     paste0(LIG,"_corr_allgenes_",Z,".RData"))
      load(paste0("~/Dropbox/GDB_archive/CMapCorr_files/",temp))
      
      temp_ds <- paste(
        switch(LIG,
               lig16="16 ligands in 14 cell lines,",
               lig295="295 ligands in 9 cell lines,"),
        switch(DAT,
               lvl4_data="landmark genes only.",
               lvl4_data_all="all genes (including inferred)."))
      
      if (Z %in% c("ligct","rep")) {
        temp_lig <- as.factor(sub("_[A-Za-z0-9_]+$","",names(CORS)))
        temp_ord1 <- tapply(CORS,temp_lig,function(X) names(X)[order(sapply(X,median))])
        temp_ord2 <- order(tapply(CORS,temp_lig,function(X) median(unlist(X))))
        temp_order <- temp_ord1[temp_ord2]
        CORS <- CORS[unlist(temp_order)]
        temp_border <- qualitative_hcl(length(temp_order),palette="dark3")
        temp_border <- unlist(mapply(rep,
                                     x=temp_border,
                                     times=sapply(temp_order,length),
                                     SIMPLIFY=F))
        temp_col <- alpha(temp_border,0.5)
        temp_bot <- switch(1 + (length(temp_order) < 20),1.5,4)
      } else {
        temp_order <- order(sapply(CORS,median))
        CORS <- CORS[temp_order]
        temp_border <- qualitative_hcl(length(CORS),palette="dark3")
        temp_col <- qualitative_hcl(length(CORS),palette="dark3",alpha=0.5)
        temp_bot <- switch(1 + (length(CORS) < 20),1.5,4)
      }
      if (length(CORS) > 300) {
        temp_border <- alpha(temp_border,0.5)
        temp_col <- alpha(temp_col,0.2)
      }      
      par(mar=c(temp_bot,4,2,1),mgp=2:0)
      boxplot(CORS,outline=T,pch=".",xaxt="n",yaxs="i",ylim=c(-1,1),
              border=temp_border,col=temp_col,
              ylab=paste("Pairwise Spearman correlation of",
                         paste("Z-scores per",
                               switch(Z,
                                      lig="ligand",
                                      ct="cell line",
                                      ligct="ligand & cell line",
                                      rep="treatment condition")),
                         sep="\n"),
              main=temp_ds)
      if (length(CORS) > 300) {
        points(sapply(CORS,median),pch=".",col=alpha(temp_border,1))
      }
      abline(h=0,col="red",lty=2)
      if (temp_bot > 1.5) {
        if (Z %in% c("ligct","rep")) {
          mtext(names(temp_order),side=1,
                at=cumsum(sapply(temp_order,length)) - sapply(temp_order,length) / 2,
                las=2,xpd=NA,line=0.1,
                col=alpha(temp_border,1)[cumsum(sapply(temp_order,length))])
        } else {
          mtext(names(CORS),side=1,at=seq_along(CORS),
                las=2,xpd=NA,line=0.1,col=alpha(temp_border,1))
        }
      } else {
        mtext(switch(Z,
                     lig="Ligands",
                     ct="Cell lines",
                     ligct="Ligands per cell line",
                     rep="Treatment conditions"),
              side=1,line=0.2,cex=1.1,font=2)
      }
    }
  }
}
```

Datasets tested, in order:
- 16 ligands in 14 cell lines, landmark genes only.
- 16 ligands in 14 cell lines, all genes (including inferred).
- 295 ligands in 9 cell lines, landmark genes only.
- 295 ligands in 9 cell lines, all genes (including inferred).


# Z-score correlation per cell line
```{r ct, fig.height=4,fig.width=9}
CORR_BXP("ct")
```


# Z-score correlation per ligand
```{r lig, fig.height=4,fig.width=9}
CORR_BXP("lig")
```


# Z-score correlation per ligand & cell line
```{r ligct, fig.height=4,fig.width=9}
CORR_BXP("ligct")
```


# Z-score correlation per treatment condition (replicate)
```{r rep, fig.height=4,fig.width=9}
CORR_BXP("rep")
```



# Comparison of all correlation distributions

```{r compare_all, fig.height=8,fig.width=4}
for (LIG in c("lig16","lig295")) {
  for (DAT in c("lvl4_data","lvl4_data_all")) {
    par(mfrow=c(4,1),mar=c(3,3,2,1),mgp=2:0)
    for (Z in c("ct","lig","ligct","rep")) {
      temp <- switch(grepl("all",DAT) + 1,
                     paste0(LIG,"_corr_",Z,".RData"),
                     paste0(LIG,"_corr_allgenes_",Z,".RData"))
      load(paste0("~/Dropbox/GDB_archive/CMapCorr_files/",temp))
      if (exists("temp_cors")) { 
        temp_w <- wilcox.test(unlist(CORS),temp_cors)
        temp_d <- median(unlist(CORS)) - median(temp_cors)
       }
      temp_cors <- unlist(CORS)
      hist(temp_cors,xlim=c(-1,1),main=NA,
           xlab=paste("Pairwise Spearman correlation of Z-scores per",
                      switch(Z,
                             lig="ligand",
                             ct="cell line",
                             ligct="ligand & cell line",
                             rep="treatment condition")))
      abline(v=median(temp_cors),col="red")
      if (Z == "ct") { 
        title(main=paste(
          switch(LIG,
                 lig16="16 ligands in 14 cell lines,",
                 lig295="295 ligands in 9 cell lines,"),
          switch(DAT,
                 lvl4_data="landmark genes only.",
                 lvl4_data_all="all genes (including inferred).")
        )) 
      }
      if (exists("temp_w")) {
        legend("topleft",bty="n",lty=c(1,NA,NA),col=c("red",NA,NA),
               legend=c("Median",
                        paste("Diff =",signif(temp_d,2)),
                        paste("p =",signif(temp_w$p.value,3))))
      } else {
        legend("topleft",bty="n",lty=1,col="red",legend="Median")
      }
      rm(temp_w,temp_d)
    }
    rm(temp_cors)
  }
}

```

```{r}
rm(list=ls())
```

```{r}
if (!require(cmapR)) { stop("install cmapR") }
load("~/Dropbox/GDB_archive/CMapCorr_files/lvl4_inputs.RData")
load("~/Dropbox/GDB_archive/CMapCorr_files/lig16_corr_lig.RData")
```

# New stuff

### Z-score fat tails?

```{r qq_zscore, fig.height=4,fig.width=8}
temp_col <- sequential_hcl(4,palette="Red-Blue")
par(mar=c(3,3,2,1),mgp=2:0)
qqnorm(as.vector(lvl4_data@mat[,lvl4_data@cdesc$pert_iname == "GDNF"]),
       main="QQ plots",pch=".",cex=2,col=temp_col[1])
points(qqnorm(as.vector(lvl4_data@mat[,lvl4_data@cdesc$pert_iname == "IGF1"]),plot.it=F),
       col=temp_col[2],pch=".",cex=2)
points(qqnorm(as.vector(lvl4_data@mat[,lvl4_data@cdesc$pert_iname == "IFNG"]),plot.it=F),
       col=temp_col[3],pch=".",cex=2)
points(qqnorm(as.vector(lvl4_data@mat[,lvl4_data@cdesc$pert_iname == "TNF"]),plot.it=F),
       col=temp_col[4],pch=".",cex=2)

```
Maybe...?


## Better correlation figure

```{r load_all_cors}
rm(list=ls())
source("~/Dropbox/GDB/line2user.R")
temp <- load("~/Dropbox/GDB_archive/CMapCorr_files/lig295_corr_lig.RData")
cor_lig <- get(temp)
temp <- load("~/Dropbox/GDB_archive/CMapCorr_files/lig295_corr_ligct.RData")
cor_ligct <- get(temp)
temp <- load("~/Dropbox/GDB_archive/CMapCorr_files/lig295_corr_rep.RData")
cor_rep <- get(temp)
rm(list=c(temp,"temp"))
```

```{r cors_per_lig, fig.height=4,fig.width=9}
cor_ligct_lig <- sapply(strsplit(names(cor_ligct),"_"),"[[",1)
cor_rep_lig <- sapply(strsplit(names(cor_rep),"_"),"[[",1)

# boxplot(list(LIG=cor_lig[["AGT"]],
#              LIGCT=unlist(cor_ligct[cor_ligct_lig == "AGT"]),
#              REP=unlist(cor_rep[cor_rep_lig == "AGT"])))

diff_ligct <- sapply(names(cor_lig),function(LIG) 
  mean(unlist(cor_ligct[cor_ligct_lig == LIG])) - mean(cor_lig[[LIG]]))
mean_ligct <- sapply(names(cor_lig),function(LIG) 
  mean(c(unlist(cor_ligct[cor_ligct_lig == LIG]),cor_lig[[LIG]])))

diff_rep <- sapply(names(cor_lig),function(LIG) 
  mean(unlist(cor_rep[cor_rep_lig == LIG])) - mean(cor_lig[[LIG]]))
mean_rep <- sapply(names(cor_lig),function(LIG) 
  mean(c(unlist(cor_rep[cor_rep_lig == LIG]),cor_lig[[LIG]])))

temp_labels <- c("ADM","IL1A","TNF")

layout(rbind(1:3),widths=c(2,3,1))
par(mar=c(5.5,4,2,1),mgp=2:0)
boxplot(unlist(sapply(temp_labels,function(X) 
  list(Lig=cor_lig[[X]],
       LigCT=unlist(cor_ligct[cor_ligct_lig == X]),
       Rep=unlist(cor_rep[cor_rep_lig == X])),
  simplify=F),recursive=F),xaxt="n",
  col=as.vector(sapply(qualitative_hcl(length(temp_labels),palette="dark3"),rep,3)))
mtext(temp_labels,side=3,line=0.1,
      col=qualitative_hcl(length(temp_labels),palette="dark3"),
      at=seq(2,by=3,length.out=length(temp_labels)))
mtext("Pairwise Spearman correlation of Z-scores",side=2,line=2.2)
text(seq(1,length(temp_labels) * 3),line2user(0.5,1),
     rep(c("By ligand","By ligand / cell line","By replicate"),length(temp_labels)),
     xpd=NA,srt=40,adj=1)
abline(v=seq(3.5,by=3,length.out=length(temp_labels) - 1),lty=2)

par(mar=c(4,4,1,0))
plot(NA,NA,xlab=NA,ylab=NA,
     xlim=range(c(mean_ligct,mean_rep)),ylim=range(diff_ligct,diff_rep))
mtext("Mean pairwise Spearman correlation of Z-scores",side=1,line=2.2)
mtext("Difference in mean vs. by ligand",side=2,line=2.2)
abline(h=0,lty=2)
points(mean_ligct,diff_ligct,pch=20,col=alpha("red",0.5))
text(mean_ligct[temp_labels],diff_ligct[temp_labels],
     temp_labels,col="red",pos=c(2,4,4),offset=0.2)
points(mean_rep,diff_rep,pch=20,col=alpha("dodgerblue",0.5))
text(mean_rep[temp_labels],diff_rep[temp_labels],
     temp_labels,col="dodgerblue",pos=c(2,4,4),offset=0.2)
legend("topright",c("By ligand / cell line","By replicate"),
       pch=20,col=alpha(c("red","dodgerblue"),0.5))

par(mar=c(4,0,1,4))
boxplot(list(diff_ligct,diff_rep),col=c("red","dodgerblue"),
        yaxt="n",xaxt="n")
abline(h=0,lty=2)
axis(4)
mtext("Difference in mean vs. by ligand",side=4,line=2.2)
if (wilcox.test(diff_ligct)$p.value < 2.2e-16) {
  mtext("*",side=3,at=1,line=-0.5)
}
if (wilcox.test(diff_rep)$p.value < 2.2e-16) {
  mtext("*",side=3,at=2,line=-0.5)
}

```

The * denotes p < 2.2e-16 from wilcoxon signed-rank test vs. mu of 0.


